---
title: "Experiment 1"
author: "Hiromichi Hagihara"
date: "2025-07-27"
output: html_document
---

```{r, results = FALSE}
library(tidyverse)
library(rlang)
library(ggrepel)
library(ggExtra)
library(purrr)
library(here)
library(scales)
library(philentropy)
library(car)
library(MASS)
library(Rtsne)

# Reference about color palette: https://r-graphics.org/recipe-colors-palette-discrete
```

The models used (10)
- gpt-4o-2024-08-06
- gpt-4o-2024-05-13
- gpt-4o-mini-2024-07-18
- meta-llama_Meta-Llama-3-70B-Instruct
- meta-llama_Meta-Llama-3-8B-Instruct
- meta-llama_Meta-Llama-3.1-405B-Instruct
- meta-llama_Meta-Llama-3.1-70B-Instruct
- meta-llama_Meta-Llama-3.1-8B-Instruct
- mistralai_Mixtral-8x22B-Instruct-v0.1
- microsoft_WizardLM-2-8x22B
- Qwen_Qwen2-72B-Instruct

Representative models (4)
- gpt-4o-2024-05-13
- gpt-4o-mini-2024-07-18
- meta-llama_Meta-Llama-3.1-405B-Instruct
- meta-llama_Meta-Llama-3.1-70B-Instruct

Word features (21)
- concreteness
- iconicity
- childBOI
- adultBOI
- Auditory
- Gustatory
- Haptic
- Interoceptive
- Olfactory
- Visual
- Foot_leg
- Hand_arm
- Head
- Mouth
- Torso
- socialness
- valence
- arousal
- dominance
- imageability
- babiness
* Excluded: phoneme, semD, frequency, mlu, contextD

# Preparation

```{r}
d <- read.csv("dataset/exp1/concatenated_results.csv", header=TRUE) %>% dplyr::select(seq1:word_used, contains("_average")) %>% 
  rename_with(~ gsub("_average", "", .), contains("_average")) %>% 
  mutate(valence=-valence+10,
         arousal=-arousal+10) %>% 
  mutate(category2=case_when(category2=="verbs" | category2=="adjectives"~"predicates", TRUE~category2)) %>% 
  dplyr::select(-phoneme, -semD, -frequency, -mlu, -contextD)

d.head <- d %>% dplyr::select(seq1:note)

d.human <- d %>% dplyr::select(seq1:babiness) %>% mutate(RatedBy = "human")

tmp <- "gpt.4o.2024.05.13_"
d.gpt1 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 
  
tmp <- "gpt.4o.mini.2024.07.18_"
d.gpt2 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 

tmp <- "meta.llama_Meta.Llama.3.1.405B.Instruct_"
d.meta1 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 

tmp <- "meta.llama_Meta.Llama.3.1.70B.Instruct_"
d.meta2 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp)

d$word_used %>% unique() %>% length() # A total of 695 words

tmp <- d.human %>% filter(category2=="function_words")
tmp <- d.gpt1 %>% filter(category2=="function_words")

rm(tmp)
```


# Correlations

## Overall

```{r}
cols <- d.human %>% dplyr::select(concreteness:babiness) %>% names()

cor.gpt1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.gpt1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o`=cor)

cor.gpt2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.gpt2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o-mini`=cor)

cor.meta1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.meta1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-405B`=cor)

cor.meta2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.meta2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-70B`=cor)

cor.all.res <- cor.gpt1.res %>% left_join(cor.gpt2.res, by="wordFeatures") %>% 
  left_join(cor.meta1.res, by="wordFeatures") %>% left_join(cor.meta2.res, by="wordFeatures")

rm(cor.gpt1.res, cor.gpt2.res, cor.meta1.res, cor.meta2.res)
```


```{r}
cor.all.res.long <- cor.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

cor.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(cor.all.res.long$cor), max(cor.all.res.long$cor)), breaks=seq(-1,1,0.2),
                       low="white", mid="yellow", high="#E60012", midpoint=0.5,
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "Cor_All.png"), dpi=350, width=6, height=9)
```


## By syntactic category

```{r}
cats <- d.head$category2 %>% unique()

cor.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o`=cor) # errors due to impossible calculation of correlation coefficients for some features (sd=0)

cor.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o-mini`=cor)

cor.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-405B`=cor)

cor.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-70B`=cor)

cor.all.cat.res <- cor.gpt1.res %>% left_join(cor.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(cor.meta1.res, by=c("category", "wordFeatures")) %>% left_join(cor.meta2.res, by=c("category", "wordFeatures"))

rm(cor.gpt1.res, cor.gpt2.res, cor.meta1.res, cor.meta2.res)
```


```{r}
cor.all.cat.res.long <- cor.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

cor.all.cat.res.long %>% subset(category=="function_words") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(cor.all.cat.res.long$cor), max(cor.all.cat.res.long$cor)), breaks=seq(-1,1,0.2),
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       low="white", mid="yellow", high="#E60012", midpoint=0.5,
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  cor.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(cor.all.cat.res.long$cor), max(cor.all.cat.res.long$cor)), breaks=seq(-1,1,0.2),
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       low="white", mid="yellow", high="#E60012", midpoint=0.5,
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("Cor_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# KL divergence

## Example

```{r}
d1 <- d.human$iconicity
d2 <- d.gpt1$iconicity

breaks <- seq(floor(min(c(d1, d2), na.rm=TRUE)), max(c(d1, d2), na.rm=TRUE)+0.2, by=0.2)

# ヒストグラムを計算（確率密度）
hist1 <- hist(d1, breaks=breaks, plot=FALSE)
hist2 <- hist(d2, breaks=breaks, plot=FALSE)

# 確率密度を計算
prob1 <- hist1$counts / sum(hist1$counts)
prob2 <- hist2$counts / sum(hist2$counts)

# philentropy packageでKL距離を計算
KL(rbind(prob1, prob2), epsilon=1e-10) # ビンの値が0の場合Infになるため小さな値を足しておく
KL(rbind(prob2, prob1), epsilon=1e-10) # ビンの値が0の場合Infになるため小さな値を足しておく

# 自前で計算
prob1 <- prob1 + 1e-10  
prob2 <- prob2 + 1e-10

sum(prob1 * log2(prob1 / prob2), na.rm=TRUE)
sum(prob2 * log2(prob2 / prob1), na.rm=TRUE)

rm(d1, d2, breaks, hist1, hist2, prob1, prob2)
```


## Overall

```{r}
calc_kl <- function(df1, df2, name){
  d1 <- df1[[name]]
  d2 <- df2[[name]]
  breaks <- seq(floor(min(c(d1, d2), na.rm=TRUE)), max(c(d1, d2), na.rm=TRUE)+0.2, by=0.2)

  hist1 <- hist(d1, breaks=breaks, plot=FALSE)
  hist2 <- hist(d2, breaks=breaks, plot=FALSE)

  prob1 <- hist1$counts / sum(hist1$counts)
  prob2 <- hist2$counts / sum(hist2$counts)

  kl <- suppressMessages(KL(rbind(prob1, prob2), epsilon=1e-10))
  return(as.numeric(kl))
}

calc_kl(d.human, d.gpt1, "iconicity")
```


```{r}
cols <- d.human %>% dplyr::select(concreteness:babiness) %>% names()

kld.gpt1.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.gpt1, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Gpt-4o`=kl)

kld.gpt2.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.gpt2, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Gpt-4o-mini`=kl)

kld.meta1.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.meta1, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Llama-3.1-405B`=kl)

kld.meta2.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.meta2, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Llama-3.1-70B`=kl)

kld.all.res <- kld.gpt1.res %>% left_join(kld.gpt2.res, by="wordFeatures") %>% 
  left_join(kld.meta1.res, by="wordFeatures") %>% left_join(kld.meta2.res, by="wordFeatures")

rm(kld.gpt1.res, kld.gpt2.res, kld.meta1.res, kld.meta2.res)
```


```{r}
kld.all.res.long <- kld.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="kld") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso"))) 

kld.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=kld))+
  geom_text(aes(label=number(kld, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(kld.all.res.long$kld), max(kld.all.res.long$kld)), breaks=seq(0,30,4),
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       low="white", mid="yellow", high="#E60012", midpoint=8,
                       name="KL divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "KLD_All.png"), dpi=350, width=6, height=9)
```


## By syntactic category

```{r}
kld.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, kl=kl_value)})}) %>% 
  rename(`Gpt-4o`=kl)

kld.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, kl=kl_value)})}) %>% 
  rename(`Gpt-4o-mini`=kl)

kld.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, kl=kl_value)})}) %>% 
  rename(`Llama-3.1-405B`=kl)

kld.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, kl=kl_value)})}) %>% 
  rename(`Llama-3.1-70B`=kl)

kld.all.cat.res <- kld.gpt1.res %>% left_join(kld.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(kld.meta1.res, by=c("category", "wordFeatures")) %>% left_join(kld.meta2.res, by=c("category", "wordFeatures"))

rm(kld.gpt1.res, kld.gpt2.res, kld.meta1.res, kld.meta2.res)
```


```{r}
kld.all.cat.res.long <- kld.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="kld") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

kld.all.cat.res.long %>% subset(category=="function_words") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=kld))+
  geom_text(aes(label=number(kld, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(kld.all.cat.res.long$kld), max(kld.all.cat.res.long$kld)), breaks=seq(0,30,5),
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       low="white", mid="yellow", high="#E60012", midpoint=15,
                       name="KL divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  kld.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=kld))+
  geom_text(aes(label=number(kld, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(kld.all.cat.res.long$kld), max(kld.all.cat.res.long$kld)), breaks=seq(0,30,5),
                       # low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       low="white", mid="yellow", high="#E60012", midpoint=15,
                       name="KL divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("KLD_Cat", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# Raw histogram and scatter plot

## Example

```{r}
i <- 1
col_name <- cols[i]
d.tmp <- d.gpt1
model_name <- "Gpt-4o"

tmp <- d.human %>% dplyr::select(-AoA) %>% rbind(d.tmp) %>% dplyr::select(seq1:note, col_name, RatedBy) %>% 
  mutate(RatedBy=case_when(RatedBy=="human"~"Human", TRUE~"Model"),
         Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  pivot_wider(names_from=RatedBy, values_from=col_name) %>% 
  filter(!is.na(Human) & !is.na(Model))

gp <- tmp %>% ggplot(aes(x=Human, y=Model, color=Category))+
  geom_jitter(size=2, width=0.005, height=0.005, alpha=0.7)+
  theme_bw()+
  labs(title=col_name, y=model_name)+
  theme(plot.title=element_text(size=20, hjust=0.5, vjust=0),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.title=element_text(size=10, color="black", hjust=0.2),
        legend.text=element_text(size=8, color="black"),
        legend.position=c(0,1),
        legend.justification=c(0,1),
        legend.background=element_rect(fill=NA))

gp <- ggMarginal(gp, type="histogram", margins="both", size=5, groupColour=TRUE, groupFill=TRUE, bins=10)
ggsave(gp, file=here("figures", "exp1", model_name, paste0(model_name, "_", i, "_", col_name, ".png")), dpi=350, width=4, height=4.5)

rm(tmp, d.tmp, gp, col_name, i, model_name)
```


## Overall

```{r}
plot_scatHist <- function(i, d.tmp, model_name){
  col_name <- cols[i]
  tmp <- d.human %>% dplyr::select(-AoA) %>% rbind(d.tmp) %>% dplyr::select(seq1:note, col_name, RatedBy) %>% 
    mutate(RatedBy=case_when(RatedBy=="human"~"Human", TRUE~"Model"),
           Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
    pivot_wider(names_from=RatedBy, values_from=col_name) %>% 
    filter(!is.na(Human) & !is.na(Model))
  
  gp <- tmp %>% ggplot(aes(x=Human, y=Model, color=Category))+
    geom_jitter(size=2, width=0.005, height=0.005, alpha=0.7)+
    theme_bw()+
    labs(title=col_name, y=model_name)+
    theme(plot.title=element_text(size=20, hjust=0.5, vjust=0),
          axis.ticks=element_line(color="black"),
          axis.text=element_text(size=14, color="black"),
          axis.title=element_text(size=14, color="black"),
          axis.text.x=element_text(size=14, color="black"),
          strip.text=element_text(size=14),
          legend.title=element_text(size=10, color="black", hjust=0.2),
          legend.text=element_text(size=8, color="black"),
          legend.position=c(0,1),
          legend.justification=c(0,1),
          legend.background=element_rect(fill=NA))
  
  gp <- ggMarginal(gp, type="histogram", margins="both", size=5, groupColour=TRUE, groupFill=TRUE, bins=10)
  ggsave(gp, file=here("figures", "exp1", model_name, paste0(model_name, "_", i, "_", col_name, ".png")), dpi=350, width=4, height=4.5)
}

plot_scatHist(2, d.gpt1, "Gpt-4o")
```


```{r}
for(i in 1:length(cols)){plot_scatHist(i, d.gpt1, "Gpt-4o")}
for(i in 1:length(cols)){plot_scatHist(i, d.gpt2, "Gpt-4o-mini")}
for(i in 1:length(cols)){plot_scatHist(i, d.meta1, "Llama-3.1-405B")}
for(i in 1:length(cols)){plot_scatHist(i, d.meta2, "Llama-3.1-70B")}
```



# Missing proportions

## Overall

```{r}
d$word_used %>% unique() %>% length() # A total of 695 words

mis.human.res <- d.human %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  # summarize(across(everything(), ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(Human=misCount)

mis.gpt1.res <- d.gpt1 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o`=misCount)

mis.gpt2.res <- d.gpt2 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o-mini`=misCount)

mis.meta1.res <- d.meta1 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-405B`=misCount)

mis.meta2.res <- d.meta2 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-70B`=misCount)

mis.all.res <- mis.human.res %>% left_join(mis.gpt1.res, by="wordFeatures") %>% left_join(mis.gpt2.res, by="wordFeatures") %>% 
  left_join(mis.meta1.res, by="wordFeatures") %>% left_join(mis.meta2.res, by="wordFeatures")

rm(mis.human.res, mis.gpt1.res, mis.gpt2.res, mis.meta1.res, mis.meta2.res)
```


```{r}
mis.all.res.long <- mis.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="misCount") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

mis.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.res.long$misCount), max(mis.all.res.long$misCount)), breaks=seq(0,510,100),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=250,
                       name="#Missing items")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "Mis_All.png"), dpi=350, width=6, height=9)
```


```{r}
mis.human.res2 <- d.human %>% pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="value") %>% 
  mutate(countFeature=case_when(is.na(value)~0, TRUE~1),
         Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  group_by(seq1, item_id, category1, Category, item_definition, word_1) %>% 
  summarize(countFeature=sum(countFeature)) %>% ungroup()

mis.human.res2 %>% ggplot(aes(x=countFeature, fill=Category))+
  geom_histogram(aes(y=after_stat(count/sum(count))), bins=22, color="white")+
  scale_x_continuous(breaks=seq(0,24,4))+
  # scale_y_continuous(limits=c(0, 300), breaks=seq(0,300,40))+
  theme_bw()+
  labs(x="# Word features", y="# Word items")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

mis.human.res2 %>% group_by(countFeature) %>% 
  summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00    2 3.00    8    9 12.00 13.00   14 15.00 16.00 17.00 18.00  19.00  20.00 21.00
# N.word       32.00 8.00    3 4.00    1    1  5.00 10.00    3 10.00 34.00 80.00 25.00 123.00 283.00 73.00
# Prop          0.05 0.01    0 0.01    0    0  0.01  0.01    0  0.01  0.05  0.12  0.04   0.18   0.41  0.11

mis.human.res2 %>% subset(Category=="nouns") %>% group_by(countFeature) %>% 
  summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature 0.00 1.00    3    8    9 12.00 13.00 15.00 16.00 17.00 18.00 19.00  20.00 21.00
# N.word       2.00 5.00    1    1    1  4.00  8.00  5.00  3.00  4.00 20.00 88.00 154.00 20.00
# Prop         0.01 0.02    0    0    0  0.01  0.03  0.02  0.01  0.01  0.06  0.28   0.49  0.06

mis.human.res2 %>% subset(Category=="predicates") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature 0.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       1.00  1.00  5.00  3.00 19.00 92.00 45.00
# Prop         0.01  0.01  0.03  0.02  0.11  0.55  0.27

mis.human.res2 %>% subset(Category=="function_words") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00 2.00 3.00 13.00 14.00 15.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       12.00 1.00 1.00 3.00  1.00  1.00  2.00 19.00 62.00  1.00  1.00  5.00  2.00
# Prop          0.11 0.01 0.01 0.03  0.01  0.01  0.02  0.17  0.56  0.01  0.01  0.05  0.02

mis.human.res2 %>% subset(Category=="other") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00 2.00 12.00 13.00 14.00 15.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       17.00 2.00 2.00  1.00  1.00  2.00  3.00 11.00  9.00  1.00 15.00 32.00  6.00
# Prop          0.17 0.02 0.02  0.01  0.01  0.02  0.03  0.11  0.09  0.01  0.15  0.31  0.06
```


## By syntactic category

```{r}
d %>% group_by(category2) %>% dplyr::select(category2, word_used) %>% unique() %>% summarize(N=n()) 
# function_words   111
# nouns            316
# other            102
# predicates       166

mis.human.res <- d.human %>% group_by(category2) %>% 
  # summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>% 
  summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(Human=misCount)

mis.gpt1.res <- d.gpt1 %>% group_by(category2) %>% 
  # summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>% 
  summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o`=misCount)

mis.gpt2.res <- d.gpt2 %>% group_by(category2) %>% 
  # summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>% 
  summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o-mini`=misCount)

mis.meta1.res <- d.meta1 %>% group_by(category2) %>% 
  # summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>% 
  summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-405B`=misCount)

mis.meta2.res <- d.meta2 %>% group_by(category2) %>% 
  # summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>% 
  summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-70B`=misCount)

mis.all.cat.res <- mis.human.res %>% left_join(mis.gpt1.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.gpt2.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.meta1.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.meta2.res, by=c("category2", "wordFeatures")) %>% 
  rename(category=category2)

rm(mis.human.res, mis.gpt1.res, mis.gpt2.res, mis.meta1.res, mis.meta2.res)
```


```{r}
mis.all.cat.res.long <- mis.all.cat.res %>% pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="misCount") %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other")),
         wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

mis.all.cat.res.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.cat.res.long$misCount), max(mis.all.cat.res.long$misCount)), breaks=seq(0,100,20),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=20,
                       name="%Missing")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  mis.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.cat.res.long$misCount), max(mis.all.cat.res.long$misCount)), breaks=seq(0,100,20),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=20,
                       name="%Missing")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("Mis_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# Relationships between Correlation and KL divergence

## Overall

```{r}
all.res.long <- cor.all.res.long %>% left_join(kld.all.res.long, by=c("wordFeatures", "model")) %>% 
  mutate(kld.log=log(kld))

all.res.long %>% ggplot(aes(x=cor, y=kld.log, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(0.1,1), breaks=seq(0,1,.2))+
  theme_bw()+
  labs(x="Correlation coefficient", y="KL divergence (log)")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
ggsave(file=here("figures", "exp1", "Scat_All.png"), dpi=350, width=11, height=9.5)   

# Correlation coefficient
models <- all.res.long$model %>% unique()

map_dfr(models, function(model_name) {
  res <- cor.test(all.res.long[all.res.long$model==model_name,]$cor, 
                  all.res.long[all.res.long$model==model_name,]$kld.log, method="pearson", use="complete.obs")
  data.frame(model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})
#                  model        cor       p.val        lwr          upr
# cor...1         Gpt-4o -0.5851166 0.005331362 -0.8117626 -0.205281268
# cor...2    Gpt-4o-mini -0.2706983 0.235284597 -0.6288944  0.182290372
# cor...3 Llama-3.1-405B -0.2743678 0.228748570 -0.6312846  0.178455408
# cor...4  Llama-3.1-70B -0.4246370 0.055020297 -0.7236680  0.008632162
```


## By syntactic category

```{r}
all.res.cat.long <- cor.all.cat.res.long %>% left_join(kld.all.cat.res.long, by=c("category", "wordFeatures", "model")) %>% 
  mutate(kld.log=log(kld))

all.res.cat.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=cor, y=kld.log, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(0,1), breaks=seq(0,1,.2))+
  theme_bw()+
  labs(title="Nouns", x="Correlation coefficient", y="KL divergence (log)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

map_dfr(cats, function(cat_name) {
  all.res.cat.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=cor, y=kld.log, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.2,1), breaks=seq(-0.2,1,.2))+
  theme_bw()+
  labs(title=cat_name, x="Correlation coefficient", y="KL divergence (log)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none") -> gp
  ggsave(file=here("figures", "exp1", paste0("Scat_Cat_", cat_name, ".png")), dpi=350, width=11, height=9.5)})

# Correlation coefficient
map_dfr(cats, function(cat_name){
  df <- all.res.cat.long %>% subset(category==cat_name)
  map_dfr(models, function(model_name) {
  res <- cor.test(df[df$model==model_name,]$cor, 
                  df[df$model==model_name,]$kld.log, method="pearson", use="complete.obs")
  data.frame(category=cat_name, model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})})
#                category          model        cor       p.val        lwr         upr
# cor...1           other         Gpt-4o -0.3478852 0.122268250 -0.6777841  0.09861067
# cor...2           other    Gpt-4o-mini -0.3650647 0.103688236 -0.6882826  0.07908539
# cor...3           other Llama-3.1-405B -0.1879146 0.414667326 -0.5731105  0.26529284
# cor...4           other  Llama-3.1-70B -0.3201315 0.157141668 -0.6605347  0.12944403
# cor...5           nouns         Gpt-4o -0.1846008 0.423092596 -0.5708006  0.26848117
# cor...6           nouns    Gpt-4o-mini -0.2078466 0.365953183 -0.5868762  0.24590377
# cor...7           nouns Llama-3.1-405B -0.3443048 0.126423131 -0.6755790  0.10263711
# cor...8           nouns  Llama-3.1-70B -0.1151133 0.619279875 -0.5209144  0.33312754
# cor...9      predicates         Gpt-4o -0.4850352 0.025837854 -0.7580078 -0.06747701
# cor...10     predicates    Gpt-4o-mini -0.2313481 0.312961715 -0.6028303  0.22256639
# cor...11     predicates Llama-3.1-405B -0.2167946 0.345202127 -0.5929854  0.23707987
# cor...12     predicates  Llama-3.1-70B -0.6559237 0.001244041 -0.8476074 -0.31281026
# cor...13 function_words         Gpt-4o -0.3052044 0.190699160 -0.6587543  0.15875819
# cor...14 function_words    Gpt-4o-mini -0.2468391 0.323418085 -0.6399591  0.24869056
# cor...15 function_words Llama-3.1-405B -0.3566411 0.133919606 -0.6978121  0.11642833
# cor...16 function_words  Llama-3.1-70B -0.2402977 0.321712781 -0.6261640  0.24011954
```



# AoA (correlation)

## Overall

```{r}
aoa.human.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.human[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(Human=cor)

aoa.gpt1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.gpt1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o`=cor)

aoa.gpt2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.gpt2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o-mini`=cor)

aoa.meta1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.meta1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-405B`=cor)

aoa.meta2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.meta2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-70B`=cor)

aoa.all.res <- aoa.human.res %>% left_join(aoa.gpt1.res, by="wordFeatures") %>% left_join(aoa.gpt2.res, by="wordFeatures") %>% 
  left_join(aoa.meta1.res, by="wordFeatures") %>% left_join(aoa.meta2.res, by="wordFeatures")

rm(aoa.human.res, aoa.gpt1.res, aoa.gpt2.res, aoa.meta1.res, aoa.meta2.res)
```


```{r}
aoa.all.res.long <- aoa.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

aoa.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.res.long$cor), max(aoa.all.res.long$cor)), breaks=seq(-1,1,0.2),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_All.png"), dpi=350, width=6, height=9)
```


```{r}
# Difference scores in correlations
aoa.all.diff <- aoa.all.res %>% mutate(across(`Gpt-4o`:`Llama-3.1-70B`, ~. -Human, .names="{col}")) %>% 
  dplyr::select(-Human)

aoa.all.diff.long <- aoa.all.diff %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         abs=abs(cor))

aoa.all.diff.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  # geom_text(aes(label=number(abs, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.diff.long$cor), max(aoa.all.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits=c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_All_Diff.png"), dpi=350, width=6, height=9)
```


```{r}
## Relationships between "correlation with AoA" and "Difference in correlations between Human and Model"

all.aoa.long <- aoa.all.res.long %>% subset(model!="Human") %>% rename(corAoA=cor) %>% 
  left_join(aoa.all.diff.long, by=c("wordFeatures", "model")) %>% rename(diffCor=cor, diffCorAbs=abs) %>% 
  left_join(aoa.all.res %>% dplyr::select(wordFeatures, Human), by="wordFeatures")

all.aoa.long %>% mutate(corAoAR = abs(corAoA / Human)) %>% 
  # ggplot(aes(x=corAoA, y=corAoAR, color=wordFeatures))+
  ggplot(aes(x=corAoA, y=diffCor, color=wordFeatures))+
  geom_hline(yintercept=0, color="black", linetype="dashed", size=0.5)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  # scale_x_continuous(limits=c(0.1,1), breaks=seq(0,1,.2))+
  theme_bw()+
  labs(x="Correlation with AoA", y="Difference in correlation (Human-Model)")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

all.aoa.long %>% ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
ggsave(file=here("figures", "exp1", "Scat_AoA_All.png"), dpi=350, width=11, height=9.5)   

# Correlation coefficient
map_dfr(models, function(model_name) {
  res <- cor.test(all.aoa.long[all.aoa.long$model==model_name,]$corAoA, 
                  all.aoa.long[all.aoa.long$model==model_name,]$Human, method="pearson", use="complete.obs")
  data.frame(model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})
#                  model       cor        p.val       lwr       upr
# cor...1         Gpt-4o 0.7132305 2.839279e-04 0.4067918 0.8753912
# cor...2    Gpt-4o-mini 0.6664178 9.714258e-04 0.3295320 0.8527751
# cor...3 Llama-3.1-405B 0.7314366 1.647948e-04 0.4380715 0.8839993
# cor...4  Llama-3.1-70B 0.8269629 3.793092e-06 0.6147268 0.9275313

# Regression
models
model_name <- models[1]
res <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(res)

model_name <- models[2]
res <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(res)

model_name <- models[3]
res <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(res)

model_name <- models[4]
res <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(res)
```


## By syntactic category

```{r}
aoa.human.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.human %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(Human=cor) # errors due to impossible calculation of correlation coefficients for some features (sd=0)

aoa.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o`=cor) 

aoa.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o-mini`=cor)

aoa.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-405B`=cor)

aoa.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-70B`=cor)

aoa.all.cat.res <- aoa.human.res %>% left_join(aoa.gpt1.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.meta1.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.meta2.res, by=c("category", "wordFeatures"))

rm(aoa.human.res, aoa.gpt1.res, aoa.gpt2.res, aoa.meta1.res, aoa.meta2.res)
```


```{r}
aoa.all.cat.res.long <- aoa.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

aoa.all.cat.res.long %>% subset(category=="predicates") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.res.long$cor), max(aoa.all.cat.res.long$cor)), breaks=seq(-1,1,0.1),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  aoa.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.res.long$cor), max(aoa.all.cat.res.long$cor)), breaks=seq(-1,1,0.1),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("AoA_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```


```{r}
# Difference scores in correlations
aoa.all.cat.diff <- aoa.all.cat.res %>% mutate(across(`Gpt-4o`:`Llama-3.1-70B`, ~. -Human, .names="{col}")) %>% 
  dplyr::select(-Human)

aoa.all.cat.diff.long <- aoa.all.cat.diff %>% pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         abs=abs(cor))

aoa.all.cat.diff.long %>% subset(category=="predicates") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  # geom_text(aes(label=number(abs, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.diff.long$cor), max(aoa.all.cat.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits=c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  aoa.all.cat.diff.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  # geom_text(aes(label=number(abs, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.diff.long$cor), max(aoa.all.cat.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits=c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("AoA_Cat_Diff_", cat_name, ".png")), dpi=350, width=6, height=9)})
```


```{r}
## Relationships between "correlation with AoA" and "Difference in correlations between Human and Model"

all.aoa.cat.long <- aoa.all.cat.res.long %>% subset(model!="Human") %>% rename(corAoA=cor) %>% 
  left_join(aoa.all.cat.diff.long, by=c("category", "wordFeatures", "model")) %>% rename(diffCor=cor, diffCorAbs=abs) %>% 
  left_join(aoa.all.cat.res %>% dplyr::select(category, wordFeatures, Human), by=c("category", "wordFeatures"))

all.aoa.cat.long %>% subset(category=="predicates") %>% 
  ggplot(aes(x=corAoA, y=diffCor, color=wordFeatures))+
  geom_hline(yintercept=0, color="black", linetype="dashed", size=0.5)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  # scale_x_continuous(limits=c(0.1,1), breaks=seq(0,1,.2))+
  theme_bw()+
  labs(title="Predicates", x="Correlation with AoA", y="Difference in correlation (Human-Model)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

all.aoa.cat.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(title="Nouns", x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

map_dfr(cats, function(cat_name) {
  all.aoa.cat.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(title=cat_name, x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
  ggsave(file=here("figures", "exp1", paste0("Scat_AoA_Cat_", cat_name, ".png")), dpi=350, width=11, height=9.5)})

# Correlation coefficient
map_dfr(cats, function(cat_name){
  df <- all.aoa.cat.long %>% subset(category==cat_name)
  map_dfr(models, function(model_name) {
    res <- cor.test(df[df$model==model_name,]$corAoA, 
                    df[df$model==model_name,]$Human, method="pearson", use="complete.obs")
    data.frame(category=cat_name, model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})})
#                category          model         cor        p.val         lwr       upr
# cor...1           other         Gpt-4o 0.786278904 2.378628e-05  0.53679411 0.9093190
# cor...2           other    Gpt-4o-mini 0.604008118 3.735852e-03  0.23310051 0.8214961
# cor...3           other Llama-3.1-405B 0.799343586 1.380247e-05  0.56136472 0.9152193
# cor...4           other  Llama-3.1-70B 0.718636392 2.426169e-04  0.41600516 0.8779579
# cor...5           nouns         Gpt-4o 0.660693240 1.113045e-03  0.32038399 0.8499606
# cor...6           nouns    Gpt-4o-mini 0.614619297 3.029881e-03  0.24899721 0.8269083
# cor...7           nouns Llama-3.1-405B 0.487309967 2.504885e-02  0.07044167 0.7592722
# cor...8           nouns  Llama-3.1-70B 0.683900484 6.295373e-04  0.35786709 0.8613038
# cor...9      predicates         Gpt-4o 0.695015143 4.705765e-04  0.37619888 0.8666746
# cor...10     predicates    Gpt-4o-mini 0.753171369 8.116992e-05  0.47636812 0.8941421
# cor...11     predicates Llama-3.1-405B 0.641691271 1.715049e-03  0.29046652 0.8405403
# cor...12     predicates  Llama-3.1-70B 0.749616350 9.157450e-05  0.47003152 0.8924929
# cor...13 function_words         Gpt-4o 0.007035311 9.771960e-01 -0.44860750 0.4597754
# cor...14 function_words    Gpt-4o-mini 0.202914138 4.193551e-01 -0.29157849 0.6118234
# cor...15 function_words Llama-3.1-405B 0.088441158 7.271109e-01 -0.39472742 0.5332913
# cor...16 function_words  Llama-3.1-70B 0.095281520 7.068563e-01 -0.38888788 0.5382098
```



# Clustering

## tSNE (Overall)

```{r}
d.all.int <- cor.all.res.long %>% left_join(kld.all.res.long, by=c("wordFeatures", "model")) %>% 
  left_join(all.aoa.long, by=c("wordFeatures", "model")) %>%
  mutate(corAoAR = corAoA / Human) %>% 
  dplyr::select(wordFeatures, model, cor, kld, corAoA, diffCorAbs) # alternatives: diffCor, diffCorAbs, corAoAR

# tSNE
set.seed(123)
?Rtsne
tsne.out <- Rtsne(d.all.int %>% dplyr::select(-wordFeatures, -model), pca=TRUE, perplexity=6, theta=0)

d.all.int.res <- d.all.int %>% cbind(as.data.frame(tsne.out$Y)) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

d.all.int.res %>% ggplot(aes(x=V1, y=V2, color=wordFeatures))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  geom_point(size=4)+
  geom_hline(yintercept=20, linetype="dashed")+
  geom_hline(yintercept=0, linetype="dashed")+
  geom_hline(yintercept=-25, linetype="dashed")+
  facet_grid(~model)+
  scale_y_continuous(breaks=seq(-50,50,10))+
  scale_x_continuous(breaks=seq(-50,50,10))+
  theme_bw()+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

# K-means
k.max <- 10
?kmeans
wss <- sapply(1:k.max, function(k){kmeans(tsne.out$Y, k, iter.max=15, nstart=50)$tot.withinss})
plot(1:k.max, wss, type="b", pch=19, frame=FALSE, xlab="#clusters K", ylab="Total within-clusters sum of squares")

km <- kmeans(tsne.out$Y, 4, iter.max=15, nstart=50)

d.all.int.res <- d.all.int.res %>% cbind(data.frame(cluster=as.factor(km$cluster))) %>% 
  mutate(cluster=case_match(cluster, "1"~"D", "2"~"C", "3"~"A", "4"~"B"))

d.all.int.res %>% ggplot(aes(x=V1, y=V2, color=cluster, shape=cluster))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  geom_point(size=4, alpha=0.6)+
  # geom_hline(yintercept=20, linetype="dashed")+
  # geom_hline(yintercept=0, linetype="dashed")+
  # geom_hline(yintercept=-25, linetype="dashed")+
  facet_grid(~model)+
  scale_y_continuous(breaks=seq(-50,50,10))+
  scale_x_continuous(breaks=seq(-50,50,10))+
  theme_bw()+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14))
ggsave(file=here("figures", "exp1", "tsne_All.png"), dpi=350, width=16, height=6)

d.all.int.res.long <- d.all.int.res %>% 
  # mutate(group=case_when(V2< -25~"A", V2<0~"B", V2<20~"C", TRUE~"D")) %>% 
  pivot_longer(c(cor:diffCorAbs), names_to="index", values_to="value") %>% 
  mutate(index=case_match(index, "cor"~"Correlation\n(human data)", "kld"~"KL divergence\n(human data)",
                          "corAoA"~"Correlation\n(AoA)", "diffCorAbs"~"Difference in\ncorrelation (AoA)"), 
         index=factor(index, levels=c("Correlation\n(human data)", "KL divergence\n(human data)", 
                                      "Correlation\n(AoA)", "Difference in\ncorrelation (AoA)")))

d.all.int.res.long %>% 
  mutate(model=factor(model, levels=c("Gpt-4o",  "Gpt-4o-mini",
                     "Llama-3.1-405B", "Llama-3.1-70B"))) %>% 
  ggplot(aes(x=cluster, y=value, color=wordFeatures))+
  geom_jitter(size=2, height=0, width=0.05, alpha=0.5)+
  geom_boxplot(fill="white", color="gray20", alpha=0.5)+
  # geom_hline(yintercept=0, linetype="dashed")+
  facet_grid(index~model, scales="free")+
  theme_bw()+
  labs(y="Value", x="Cluster")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
     axis.ticks=element_line(color="black"),
     axis.text=element_text(size=14, color="black"),
     axis.title=element_text(size=14, color="black"),
     axis.text.x=element_text(size=14, color="black"),
     strip.text=element_text(size=14),
     legend.position="none") 
ggsave(file=here("figures", "exp1", "tsne_box_All.png"), dpi=350, width=8, height=7.5)

d.all.int.res.long %>% dplyr::select(wordFeatures, model, cluster) %>% 
  unique() %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cluster))+
  geom_text(aes(label=cluster), size=6, color="gray30")+
  scale_fill_brewer(palette="BuGn", name="Cluster")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"),
        plot.margin=margin(0, 40, 0, 0)) # (Top, Right, Bottom, Left)
ggsave(file=here("figures", "exp1", "tsne_group_All.png"), dpi=350, width=6, height=9)
```


## PCA (Overall)

```{r}
pca <- prcomp(d.all.int %>% dplyr::select(-wordFeatures, -model), scale=TRUE)

pca.cont <- data.frame(t(summary(pca)$importance)) %>% 
  tibble::rownames_to_column(var="PCs") %>% 
  tibble::rownames_to_column(var="PCs2") %>% 
  mutate(PCs2 = as.numeric(PCs2))

pca.cont %>% ggplot(aes(x=PCs2, y=Proportion.of.Variance))+
  geom_bar(stat="identity")+
  geom_line(aes(y=Cumulative.Proportion), color="blue", linewidth=3)+
  geom_abline(slope=0, intercept=0.1, color="red")+
  theme_bw()+
  scale_x_continuous(breaks=1:20)+
  labs(x="Principal components", y="Proportion of variance")+
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16))

pca.axis <- data.frame(pca$rotation) %>% tibble::rownames_to_column(var="Column")

library(ggsci)
pca.axis %>% ggplot(aes(x=PC1, y=PC2))+
  #geom_point()+
  geom_text(aes(label=Column), hjust=0, vjust=0)+
  geom_segment(aes(x=0, y=0, xend=PC1, yend=PC2),linewidth=.4, arrow=arrow(length=unit(.1, 'cm')))+
  theme_bw()+
  scale_colour_d3()+ 
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16))

d.all.int %>% cbind(data.frame(pca$x)) %>% 
  ggplot(aes(x=PC1, y=PC2, color=wordFeatures))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  geom_point(size=4)+
  facet_grid(~model)+
  theme_bw()+
   theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none") 
```



# Other exploratory analyses

## Prediction of AoA (Overall)

```{r}
# Human
d.human.aoa <- d.human %>% dplyr::select(-socialness, -imageability) %>%  # socialness and imageability had >200 missing items
  filter(complete.cases(dplyr::select(., AoA:babiness))) %>% 
  mutate(across(concreteness:babiness, ~ scale(.)[,1], .names="{.col}")) %>% 
  mutate(Human="yes")

m.human.null <- lm(AoA ~ 1, data=d.human.aoa %>% dplyr::select(AoA, concreteness:babiness))
m.human.full <- lm(AoA ~ ., data=d.human.aoa %>% dplyr::select(AoA, concreteness:babiness))
anova(m.human.null, m.human.full)

summary(m.human.full)
vif(m.human.full)
```

```{r}
# gpt1
d.gpt1.aoa <- d.gpt1 %>% dplyr::select(-socialness, -imageability) %>%
  mutate(across(concreteness:babiness, ~ scale(.)[,1], .names="{.col}")) %>% 
  left_join(d.human %>% dplyr::select(seq1, AoA), by="seq1") %>% 
  left_join(d.human.aoa %>% dplyr::select(seq1, Human), by="seq1") %>% 
  mutate(Human=case_when(Human=="yes"~"yes", TRUE~"no")) %>% 
  filter(complete.cases(dplyr::select(., concreteness:AoA))) %>% 
  dplyr::select(-adultBOI) %>% 
  subset(Human=="yes")

m.gpt1.null <- lm(AoA ~ 1, data=d.gpt1.aoa %>% dplyr::select(AoA, concreteness:babiness))
m.gpt1.full <- lm(AoA ~ ., data=d.gpt1.aoa %>% dplyr::select(AoA, concreteness:babiness))
anova(m.gpt1.null, m.gpt1.full)

summary(m.gpt1.full)
vif(m.gpt1.full)

m.gpt1.sel <- step(m.gpt1.null, scope=list(lower=m.gpt1.null, upper=m.gpt1.full), direction="both", k=log(nrow(d.gpt1.aoa)))
summary(m.gpt1.sel)
vif(m.gpt1.sel)
```

