---
title: "Experiment 1"
author: "Hiromichi Hagihara"
date: "2025-07-27"
output: html_document
---

```{r, results = FALSE}
library(tidyverse)
library(rlang)
library(ggrepel)
library(ggExtra)
library(purrr)
library(here)
library(scales)
library(philentropy)
library(car)
library(MASS)
library(Rtsne)

# Reference about color palette: https://r-graphics.org/recipe-colors-palette-discrete
```

The models used (10)
- gpt-4o-2024-08-06
- gpt-4o-2024-05-13
- gpt-4o-mini-2024-07-18
- meta-llama_Meta-Llama-3-70B-Instruct
- meta-llama_Meta-Llama-3-8B-Instruct
- meta-llama_Meta-Llama-3.1-405B-Instruct
- meta-llama_Meta-Llama-3.1-70B-Instruct
- meta-llama_Meta-Llama-3.1-8B-Instruct
- mistralai_Mixtral-8x22B-Instruct-v0.1
- microsoft_WizardLM-2-8x22B
- Qwen_Qwen2-72B-Instruct

Representative models (4)
- gpt-4o-2024-05-13
- gpt-4o-mini-2024-07-18
- meta-llama_Meta-Llama-3.1-405B-Instruct
- meta-llama_Meta-Llama-3.1-70B-Instruct

Word features (21)
- concreteness
- iconicity
- childBOI
- adultBOI
- Auditory
- Gustatory
- Haptic
- Interoceptive
- Olfactory
- Visual
- Foot_leg
- Hand_arm
- Head
- Mouth
- Torso
- socialness
- valence
- arousal
- dominance
- imageability
- babiness
* Excluded: phoneme, semD, frequency, mlu, contextD

# Preparation

```{r}
d <- read.csv("dataset/exp1/concatenated_results.csv", header=TRUE) %>% dplyr::select(seq1:word_used, contains("_average")) %>% 
  rename_with(~ gsub("_average", "", .), contains("_average")) %>% 
  mutate(valence=-valence+10,
         arousal=-arousal+10) %>% 
  mutate(category2=case_when(category2=="verbs" | category2=="adjectives"~"predicates", TRUE~category2)) %>% 
  dplyr::select(-phoneme, -semD, -frequency, -mlu, -contextD)

d.head <- d %>% dplyr::select(seq1:note)

d.human <- d %>% dplyr::select(seq1:babiness) %>% mutate(RatedBy = "human")

tmp <- "gpt.4o.2024.05.13_"
d.gpt1 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 
  
tmp <- "gpt.4o.mini.2024.07.18_"
d.gpt2 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 

tmp <- "meta.llama_Meta.Llama.3.1.405B.Instruct_"
d.meta1 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp) 

tmp <- "meta.llama_Meta.Llama.3.1.70B.Instruct_"
d.meta2 <- d %>% dplyr::select(seq1:note, contains(tmp)) %>% rename_with(~ gsub(tmp, "", .), contains(tmp)) %>% mutate(RatedBy = tmp)

d$word_used %>% unique() %>% length() # A total of 695 words

tmp <- d.human %>% filter(category2=="function_words")
tmp <- d.gpt1 %>% filter(category2=="function_words")

rm(tmp)
```


# Correlations

## Overall

```{r}
cols <- d.human %>% dplyr::select(concreteness:babiness) %>% names()

cor.gpt1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.gpt1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o`=cor)

cor.gpt2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.gpt2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o-mini`=cor)

cor.meta1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.meta1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-405B`=cor)

cor.meta2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[[col_name]], d.meta2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-70B`=cor)

cor.all.res <- cor.gpt1.res %>% left_join(cor.gpt2.res, by="wordFeatures") %>% 
  left_join(cor.meta1.res, by="wordFeatures") %>% left_join(cor.meta2.res, by="wordFeatures")

rm(cor.gpt1.res, cor.gpt2.res, cor.meta1.res, cor.meta2.res)
```


```{r}
cor.all.res.long <- cor.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

cor.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix=""), color=cor>0.7), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(cor.all.res.long$cor), max(cor.all.res.long$cor)), breaks=seq(-1,1,0.2),
                       low="white", mid="yellow", high="#CC0000", midpoint=0.5,
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "Cor_All.png"), dpi=350, width=6, height=9)
```


## By syntactic category

```{r}
cats <- d.head$category2 %>% unique()

cor.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o`=cor) 

# errors due to impossible calculation of correlation coefficients for some features (sd=0 and/or missing data)
# function_words - Gustatory
d.gpt1 %>% subset(category2=="function_words") %>% dplyr::select(Gustatory) %>% unique()

cor.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o-mini`=cor)
# function_words - Gustatory, Olfactory, arousal

cor.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-405B`=cor)
# function_words - Gustatory, Olfactory

cor.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[[col_name]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-70B`=cor)
# function_words - Gustatory, Olfactory

cor.all.cat.res <- cor.gpt1.res %>% left_join(cor.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(cor.meta1.res, by=c("category", "wordFeatures")) %>% left_join(cor.meta2.res, by=c("category", "wordFeatures"))

rm(cor.gpt1.res, cor.gpt2.res, cor.meta1.res, cor.meta2.res)
```


```{r}
cor.all.cat.res.long <- cor.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

cor.all.cat.res.long %>% subset(category=="function_words") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix=""), color=cor>0.7), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(cor.all.cat.res.long$cor), max(cor.all.cat.res.long$cor)), breaks=seq(-1,1,0.2),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=0.5,
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  cor.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix=""), color=cor>0.7), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(cor.all.cat.res.long$cor), max(cor.all.cat.res.long$cor)), breaks=seq(-1,1,0.2),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=0.5,
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("Cor_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# KL divergence

## Example

```{r}
d1 <- d.human$iconicity
d2 <- d.gpt1$iconicity

breaks <- seq(floor(min(c(d1, d2), na.rm=TRUE)), max(c(d1, d2), na.rm=TRUE)+0.2, by=0.2)

# ヒストグラムを計算（確率密度）
hist1 <- hist(d1, breaks=breaks, plot=FALSE)
hist2 <- hist(d2, breaks=breaks, plot=FALSE)

# 確率密度を計算
prob1 <- hist1$counts / sum(hist1$counts)
prob2 <- hist2$counts / sum(hist2$counts)

# philentropy packageでKL距離を計算
KL(rbind(prob1, prob2), epsilon=1e-10) # ビンの値が0の場合Infになるため小さな値を足しておく
KL(rbind(prob2, prob1), epsilon=1e-10) # ビンの値が0の場合Infになるため小さな値を足しておく

# 自前で計算
prob1 <- prob1 + 1e-10  
prob2 <- prob2 + 1e-10

sum(prob1 * log2(prob1 / prob2), na.rm=TRUE)
sum(prob2 * log2(prob2 / prob1), na.rm=TRUE)

rm(d1, d2, breaks, hist1, hist2, prob1, prob2)
```


## Overall

```{r}
calc_kl <- function(df1, df2, name){
  d1 <- df1[[name]]
  d2 <- df2[[name]]
  breaks <- seq(floor(min(c(d1, d2), na.rm=TRUE)), max(c(d1, d2), na.rm=TRUE)+0.2, by=0.2)

  hist1 <- hist(d1, breaks=breaks, plot=FALSE)
  hist2 <- hist(d2, breaks=breaks, plot=FALSE)

  prob1 <- hist1$counts / sum(hist1$counts)
  prob2 <- hist2$counts / sum(hist2$counts)

  kl <- suppressMessages(KL(rbind(prob1, prob2), epsilon=1e-10))
  return(as.numeric(kl))
}

calc_kl(d.human, d.gpt1, "iconicity")
```


```{r}
cols <- d.human %>% dplyr::select(concreteness:babiness) %>% names()

kld.gpt1.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.gpt1, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Gpt-4o`=kl)

kld.gpt2.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.gpt2, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Gpt-4o-mini`=kl)

kld.meta1.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.meta1, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Llama-3.1-405B`=kl)

kld.meta2.res <- map_dfr(cols, function(col_name) {
  kl_value <- calc_kl(d.human, d.meta2, col_name)
  data.frame(wordFeatures=col_name, kl=kl_value)}) %>% 
  rename(`Llama-3.1-70B`=kl)

kld.all.res <- kld.gpt1.res %>% left_join(kld.gpt2.res, by="wordFeatures") %>% 
  left_join(kld.meta1.res, by="wordFeatures") %>% left_join(kld.meta2.res, by="wordFeatures")

rm(kld.gpt1.res, kld.gpt2.res, kld.meta1.res, kld.meta2.res)
```


```{r}
kld.all.res.long <- kld.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="kld") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso"))) 

kld.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=kld))+
  geom_text(aes(label=number(kld, accuracy=0.01, decimal.mark=".", suffix=""), color=kld>13), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(kld.all.res.long$kld), max(kld.all.res.long$kld)), breaks=seq(0,30,4),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=8,
                       name="KL divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
# ggsave(file=here("figures", "exp1", "KLD_All.png"), dpi=350, width=6, height=9)
```


# Jensen–Shannon divergence

```{r}
calc_js <- function(df1, df2, name, by = 0.2, eps = 1e-10, base2 = TRUE){
  d1 <- df1[[name]]; d2 <- df2[[name]]
  breaks <- seq(floor(min(c(d1, d2), na.rm=TRUE)),
                max(c(d1, d2), na.rm=TRUE) + by, by = by)
  h1 <- hist(d1, breaks = breaks, plot = FALSE)
  h2 <- hist(d2, breaks = breaks, plot = FALSE)
  p <- h1$counts; q <- h2$counts
  p <- (p + eps); p <- p / sum(p)
  q <- (q + eps); q <- q / sum(q)
  m <- 0.5 * (p + q)
  KL <- function(a, b) sum(a * log(a / b))
  js_nat <- 0.5 * KL(p, m) + 0.5 * KL(q, m)
  if (base2) js_nat / log(2) else js_nat
}

calc_js(d.human, d.gpt1, "iconicity")

# library(philentropy)
# calc_js_2 <- function(df1, df2, name, by = 0.2, eps = 1e-10){
#   d1 <- df1[[name]]; d2 <- df2[[name]]
#   breaks <- seq(floor(min(c(d1, d2), na.rm = TRUE)),
#                 max(c(d1, d2), na.rm = TRUE) + by, by = by)
#   h1 <- hist(d1, breaks = breaks, plot = FALSE)
#   h2 <- hist(d2, breaks = breaks, plot = FALSE)
#   p <- h1$counts; q <- h2$counts
#   p <- (p + eps); p <- p / sum(p)
#   q <- (q + eps); q <- q / sum(q)
#   mat <- rbind(p, q)
#   as.numeric(distance(mat, method = "jensen-shannon", unit = "log2"))
# }
# 
# calc_js(d.human, d.gpt1, "iconicity")
# calc_js_2(d.human, d.gpt1, "iconicity")
```


```{r}
cols <- d.human %>% dplyr::select(concreteness:babiness) %>% names()

jsd.gpt1.res <- map_dfr(cols, function(col_name) {
  js_value <- calc_js(d.human, d.gpt1, col_name)
  data.frame(wordFeatures=col_name, js=js_value)}) %>% 
  rename(`Gpt-4o`=js)

jsd.gpt2.res <- map_dfr(cols, function(col_name) {
  js_value <- calc_js(d.human, d.gpt2, col_name)
  data.frame(wordFeatures=col_name, js=js_value)}) %>% 
  rename(`Gpt-4o-mini`=js)

jsd.meta1.res <- map_dfr(cols, function(col_name) {
  js_value <- calc_js(d.human, d.meta1, col_name)
  data.frame(wordFeatures=col_name, js=js_value)}) %>% 
  rename(`Llama-3.1-405B`=js)

jsd.meta2.res <- map_dfr(cols, function(col_name) {
  js_value <- calc_js(d.human, d.meta2, col_name)
  data.frame(wordFeatures=col_name, js=js_value)}) %>% 
  rename(`Llama-3.1-70B`=js)

jsd.all.res <- jsd.gpt1.res %>% left_join(jsd.gpt2.res, by="wordFeatures") %>% 
  left_join(jsd.meta1.res, by="wordFeatures") %>% left_join(jsd.meta2.res, by="wordFeatures")

rm(jsd.gpt1.res, jsd.gpt2.res, jsd.meta1.res, jsd.meta2.res)
```


```{r}
jsd.all.res.long <- jsd.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="jsd") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso"))) 

jsd.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=jsd))+
  geom_text(aes(label=number(jsd, accuracy=0.01, decimal.mark=".", suffix=""), color=jsd>0.6), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(jsd.all.res.long$jsd), max(jsd.all.res.long$jsd)), breaks=seq(0,1,0.2),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=0.4,
                       name="JS divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "JSD_All.png"), dpi=350, width=6, height=9)
```


## Relationships between KL and JS divergences

```{r}
tmp <- kld.all.res.long %>% left_join(jsd.all.res.long)

tmp %>% ggplot(aes(x=kld, y=jsd, color=wordFeatures))+
  geom_point()+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  facet_wrap(~model, strip.position="top")+
  theme_bw()+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"),
        legend.position="none")  
```


## By syntactic category

```{r}
jsd.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  js_value <- calc_js(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, js=js_value)})}) %>% 
  rename(`Gpt-4o`=js)

jsd.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  js_value <- calc_js(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, js=js_value)})}) %>% 
  rename(`Gpt-4o-mini`=js)

jsd.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  js_value <- calc_js(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, js=js_value)})}) %>% 
  rename(`Llama-3.1-405B`=js)

jsd.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  js_value <- calc_js(df1, df2, col_name)
  data.frame(category=cat_name, wordFeatures=col_name, js=js_value)})}) %>% 
  rename(`Llama-3.1-70B`=js)

jsd.all.cat.res <- jsd.gpt1.res %>% left_join(jsd.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(jsd.meta1.res, by=c("category", "wordFeatures")) %>% left_join(jsd.meta2.res, by=c("category", "wordFeatures"))

rm(jsd.gpt1.res, jsd.gpt2.res, jsd.meta1.res, jsd.meta2.res)
```


```{r}
jsd.all.cat.res.long <- jsd.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="jsd") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")))

jsd.all.cat.res.long %>% subset(category=="function_words") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=jsd))+
  geom_text(aes(label=number(jsd, accuracy=0.01, decimal.mark=".", suffix=""), color=jsd>0.6), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(jsd.all.cat.res.long$jsd), max(jsd.all.cat.res.long$jsd)), breaks=seq(0,1,0.2),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=0.4,
                       name="JS divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  jsd.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=jsd))+
  geom_text(aes(label=number(jsd, accuracy=0.01, decimal.mark=".", suffix=""), color=jsd>0.6), size=6, show.legend=FALSE)+
  scale_color_manual(values=c("gray20", "white"))+
  scale_fill_gradient2(limits=c(min(jsd.all.cat.res.long$jsd), max(jsd.all.cat.res.long$jsd)), breaks=seq(0,1,0.2),
                       # low="#00A0E9", mid="white", high="#CC0000", midpoint=0, 
                       low="white", mid="yellow", high="#CC0000", midpoint=0.4,
                       name="JS divergence")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("JSD_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# Raw histogram and scatter plot

## Example

```{r}
i <- 1
col_name <- cols[i]
d.tmp <- d.gpt1
model_name <- "Gpt-4o"

tmp <- d.human %>% dplyr::select(-AoA) %>% rbind(d.tmp) %>% dplyr::select(seq1:note, col_name, RatedBy) %>% 
  mutate(RatedBy=case_when(RatedBy=="human"~"Human", TRUE~"Model"),
         Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  pivot_wider(names_from=RatedBy, values_from=col_name) %>% 
  filter(!is.na(Human) & !is.na(Model))

gp <- tmp %>% ggplot(aes(x=Human, y=Model, color=Category, shape=Category))+
  geom_jitter(size=2, width=0.005, height=0.005, alpha=0.7)+
  theme_bw()+
  labs(title=col_name, y=model_name, color=NULL, shape=NULL)+
  scale_shape_manual(values = c(16, 17, 15, 18))+
  scale_x_continuous(limits=c(min(tmp$Human, tmp$Model, na.rm=TRUE)-0.01,max(tmp$Human, tmp$Model, na.rm=TRUE)+0.1), breaks=seq(0,10,1))+
  scale_y_continuous(limits=c(min(tmp$Human, tmp$Model, na.rm=TRUE)-0.01,max(tmp$Human, tmp$Model, na.rm=TRUE)+0.1), breaks=seq(0,10,1))+
  coord_fixed(ratio=1)+
  theme(plot.title=element_text(size=25, hjust=0.5, vjust=0),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.title=element_blank(),
        legend.text=element_text(size = 9, color = "black", hjust = 0),
        legend.position="bottom",
        legend.direction="horizontal",
        legend.box="horizontal", 
        # legend.position=c(0,1),
        # legend.justification=c(0,1),
        legend.background=element_rect(fill=NA),
        legend.key = element_blank(), 
        legend.key.width = unit(3, "pt"), 
        legend.margin = margin(t = -5, b = -5, unit = "pt"),  
        legend.spacing.x = unit(8, "pt"),
        legend.spacing.y = unit(2, "pt"),     
        legend.box.margin = margin(0, 0, 0, 0, unit = "pt"))+
  guides(color = guide_legend(nrow = 1, byrow = TRUE, label.hjust = 0, keywidth = 0.5, default.unit = "cm", label.position = "right"),
         shape = guide_legend(nrow = 1, byrow = TRUE, label.hjust = 0, keywidth = 0.5, default.unit = "cm", label.position = "right"))

gp <- ggMarginal(gp, type="histogram", margins="both", size=5, groupColour=TRUE, groupFill=TRUE, bins=10)
ggsave(gp, file=here("figures", "exp1", model_name, paste0(model_name, "_", i, "_", col_name, ".png")), 
       dpi=350, width=4, height=4.5, bg="white")

rm(tmp, d.tmp, gp, col_name, i, model_name)
```


## Overall

```{r}
plot_scatHist <- function(i, d.tmp, model_name){
  col_name <- cols[i]
  tmp <- d.human %>% dplyr::select(-AoA) %>% rbind(d.tmp) %>% dplyr::select(seq1:note, col_name, RatedBy) %>% 
    mutate(RatedBy=case_when(RatedBy=="human"~"Human", TRUE~"Model"),
           Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
    pivot_wider(names_from=RatedBy, values_from=col_name) %>% 
    filter(!is.na(Human) & !is.na(Model))
  
  gp <- tmp %>% ggplot(aes(x=Human, y=Model, color=Category, shape=Category))+
    geom_jitter(size=2, width=0.005, height=0.005, alpha=0.7)+
    theme_bw()+
    labs(title=col_name, y=model_name, color=NULL, shape=NULL)+
    scale_shape_manual(values = c(16, 17, 15, 18))+
    scale_x_continuous(limits=c(min(tmp$Human, tmp$Model, na.rm=TRUE)-0.01,max(tmp$Human, tmp$Model, na.rm=TRUE)+0.1), breaks=seq(0,10,1))+
    scale_y_continuous(limits=c(min(tmp$Human, tmp$Model, na.rm=TRUE)-0.01,max(tmp$Human, tmp$Model, na.rm=TRUE)+0.1), breaks=seq(0,10,1))+
    coord_fixed(ratio=1)+
    theme(plot.title=element_text(size=25, hjust=0.5, vjust=0),
          axis.ticks=element_line(color="black"),
          axis.text=element_text(size=14, color="black"),
          axis.title=element_text(size=14, color="black"),
          axis.text.x=element_text(size=14, color="black"),
          strip.text=element_text(size=14),
          legend.title=element_blank(),
          legend.text=element_text(size = 9, color = "black", hjust = 0),
          legend.position="bottom",
          legend.direction="horizontal",
          legend.box="horizontal", 
          # legend.position=c(0,1),
          # legend.justification=c(0,1),
          legend.background=element_rect(fill=NA),
          legend.key = element_blank(), 
          legend.key.width = unit(3, "pt"), 
          legend.margin = margin(t = -5, b = -5, unit = "pt"),  
          legend.spacing.x = unit(8, "pt"),
          legend.spacing.y = unit(2, "pt"),     
          legend.box.margin = margin(0, 0, 0, 0, unit = "pt"))+
    guides(color = guide_legend(nrow = 1, byrow = TRUE, label.hjust = 0, keywidth = 0.5, default.unit = "cm", label.position = "right"),
           shape = guide_legend(nrow = 1, byrow = TRUE, label.hjust = 0, keywidth = 0.5, default.unit = "cm", label.position = "right"))
  
  gp <- ggMarginal(gp, type="histogram", margins="both", size=5, groupColour=TRUE, groupFill=TRUE, bins=10)
  ggsave(gp, file=here("figures", "exp1", model_name, paste0(model_name, "_", i, "_", col_name, ".png")), 
         dpi=350, width=4, height=4.5, bg="white")
}

plot_scatHist(2, d.gpt1, "Gpt-4o")
```


```{r}
for(i in 1:length(cols)){plot_scatHist(i, d.gpt1, "Gpt-4o")}
for(i in 1:length(cols)){plot_scatHist(i, d.gpt2, "Gpt-4o-mini")}
for(i in 1:length(cols)){plot_scatHist(i, d.meta1, "Llama-3.1-405B")}
for(i in 1:length(cols)){plot_scatHist(i, d.meta2, "Llama-3.1-70B")}
```



# Missing proportions

## Overall

```{r}
d$word_used %>% unique() %>% length() # A total of 695 words

mis.human.res <- d.human %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  # summarize(across(everything(), ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(Human=misCount)

mis.gpt1.res <- d.gpt1 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o`=misCount)

mis.gpt2.res <- d.gpt2 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o-mini`=misCount)

mis.meta1.res <- d.meta1 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-405B`=misCount)

mis.meta2.res <- d.meta2 %>% dplyr::select(concreteness:babiness) %>% 
  summarize(across(everything(), ~sum(is.na(.)))) %>% 
  pivot_longer(cols=everything(), names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-70B`=misCount)

mis.all.res <- mis.human.res %>% left_join(mis.gpt1.res, by="wordFeatures") %>% left_join(mis.gpt2.res, by="wordFeatures") %>% 
  left_join(mis.meta1.res, by="wordFeatures") %>% left_join(mis.meta2.res, by="wordFeatures")

rm(mis.human.res, mis.gpt1.res, mis.gpt2.res, mis.meta1.res, mis.meta2.res)
```


```{r}
mis.all.res.long <- mis.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="misCount") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

mis.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.res.long$misCount), max(mis.all.res.long$misCount)), breaks=seq(0,510,100),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=250,
                       name="#Missing items")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "Mis_All.png"), dpi=350, width=6, height=9)
```


```{r}
mis.human.res2 <- d.human %>% pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="value") %>% 
  mutate(countFeature=case_when(is.na(value)~0, TRUE~1),
         Category=factor(category2, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  group_by(seq1, item_id, category1, Category, item_definition, word_1) %>% 
  summarize(countFeature=sum(countFeature)) %>% ungroup()

mis.human.res2 %>% ggplot(aes(x=countFeature, fill=Category))+
  geom_histogram(aes(y=after_stat(count/sum(count))), bins=22, color="white")+
  scale_x_continuous(breaks=seq(0,24,4))+
  # scale_y_continuous(limits=c(0, 300), breaks=seq(0,300,40))+
  theme_bw()+
  labs(x="# Word features", y="# Word items")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

mis.human.res2 %>% group_by(countFeature) %>% 
  summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00    2 3.00    8    9 12.00 13.00   14 15.00 16.00 17.00 18.00  19.00  20.00 21.00
# N.word       32.00 8.00    3 4.00    1    1  5.00 10.00    3 10.00 34.00 80.00 25.00 123.00 283.00 73.00
# Prop          0.05 0.01    0 0.01    0    0  0.01  0.01    0  0.01  0.05  0.12  0.04   0.18   0.41  0.11

mis.human.res2 %>% subset(Category=="nouns") %>% group_by(countFeature) %>% 
  summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature 0.00 1.00    3    8    9 12.00 13.00 15.00 16.00 17.00 18.00 19.00  20.00 21.00
# N.word       2.00 5.00    1    1    1  4.00  8.00  5.00  3.00  4.00 20.00 88.00 154.00 20.00
# Prop         0.01 0.02    0    0    0  0.01  0.03  0.02  0.01  0.01  0.06  0.28   0.49  0.06

mis.human.res2 %>% subset(Category=="predicates") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature 0.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       1.00  1.00  5.00  3.00 19.00 92.00 45.00
# Prop         0.01  0.01  0.03  0.02  0.11  0.55  0.27

mis.human.res2 %>% subset(Category=="function_words") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00 2.00 3.00 13.00 14.00 15.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       12.00 1.00 1.00 3.00  1.00  1.00  2.00 19.00 62.00  1.00  1.00  5.00  2.00
# Prop          0.11 0.01 0.01 0.03  0.01  0.01  0.02  0.17  0.56  0.01  0.01  0.05  0.02

mis.human.res2 %>% subset(Category=="other") %>% group_by(countFeature) %>% 
    summarize(N.word=n()) %>% mutate(Prop=round(N.word/sum(N.word),2)) %>% t()
# countFeature  0.00 1.00 2.00 12.00 13.00 14.00 15.00 16.00 17.00 18.00 19.00 20.00 21.00
# N.word       17.00 2.00 2.00  1.00  1.00  2.00  3.00 11.00  9.00  1.00 15.00 32.00  6.00
# Prop          0.17 0.02 0.02  0.01  0.01  0.02  0.03  0.11  0.09  0.01  0.15  0.31  0.06
```


## By syntactic category

```{r}
d %>% group_by(category2) %>% dplyr::select(category2, word_used) %>% unique() %>% summarize(N=n()) 
# function_words   111
# nouns            316
# other            102
# predicates       166

mis.human.res <- d.human %>% group_by(category2) %>% 
  summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>%
  # summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(Human=misCount)

mis.gpt1.res <- d.gpt1 %>% group_by(category2) %>% 
  summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>%
  # summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o`=misCount)

mis.gpt2.res <- d.gpt2 %>% group_by(category2) %>% 
  summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>%
  # summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Gpt-4o-mini`=misCount)

mis.meta1.res <- d.meta1 %>% group_by(category2) %>% 
  summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>%
  # summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-405B`=misCount)

mis.meta2.res <- d.meta2 %>% group_by(category2) %>% 
  summarize(across(concreteness:babiness, ~sum(is.na(.)))) %>%
  # summarize(across(concreteness:babiness, ~round(mean(is.na(.))*100, 1))) %>%  # Calculate Prop instead of Count
  pivot_longer(cols=concreteness:babiness, names_to="wordFeatures", values_to="misCount") %>% 
  rename(`Llama-3.1-70B`=misCount)

mis.all.cat.res <- mis.human.res %>% left_join(mis.gpt1.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.gpt2.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.meta1.res, by=c("category2", "wordFeatures")) %>% 
  left_join(mis.meta2.res, by=c("category2", "wordFeatures")) %>% 
  rename(category=category2)

rm(mis.human.res, mis.gpt1.res, mis.gpt2.res, mis.meta1.res, mis.meta2.res)
```


```{r}
mis.all.cat.res.long <- mis.all.cat.res %>% pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="misCount") %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other")),
         wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

mis.all.cat.res.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.cat.res.long$misCount), max(mis.all.cat.res.long$misCount)), #breaks=seq(0,100,20),
                       low="#00A0E9", mid="white", high="#E60012", # midpoint=20,
                       # name="%Missing"
                       name="Missing")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  mis.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=misCount))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(misCount, accuracy=1, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(mis.all.cat.res.long$misCount), max(mis.all.cat.res.long$misCount)), #breaks=seq(0,100,20),
                       low="#00A0E9", mid="white", high="#E60012", #midpoint=20,
                       # name="%Missing"
                       name="Missing")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("Mis_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```



# Relationships between Correlation and JS divergence

## Overall

```{r}
all.res.long <- cor.all.res.long %>% left_join(jsd.all.res.long, by=c("wordFeatures", "model"))

all.res.long %>% ggplot(aes(x=cor, y=jsd, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.2, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits = c(0.1, 1), breaks = seq(0, 1, .2), expand = expansion(mult = c(.03, .08)))+
  scale_y_continuous(expand = expansion(mult = c(.05, .15)))+
  # scale_x_continuous(limits=c(0.1,1), breaks=seq(0,1,.2))+
  theme_bw()+
  labs(x="Correlation coefficient", y="JS divergence")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
ggsave(file=here("figures", "exp1", "Scat_All.png"), dpi=350, width=11, height=9.5)   

# Correlation coefficient
models <- all.res.long$model %>% unique()

map_dfr(models, function(model_name) {
  res <- cor.test(all.res.long[all.res.long$model==model_name,]$cor, 
                  all.res.long[all.res.long$model==model_name,]$jsd, method="pearson", use="complete.obs")
  data.frame(model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})
#                  model        cor        p.val        lwr          upr
# cor...1         Gpt-4o -0.4397022 0.0460954440 -0.7323743 -0.009893275
# cor...2    Gpt-4o-mini -0.4798936 0.0276939600 -0.7551423 -0.060802978
# cor...3 Llama-3.1-405B -0.5435386 0.0108747840 -0.7898874 -0.146142457
# cor...4  Llama-3.1-70B -0.7088185 0.0003219831 -0.8732895 -0.399318110
```



## By syntactic category

```{r}
all.res.cat.long <- cor.all.cat.res.long %>% left_join(jsd.all.cat.res.long, by=c("category", "wordFeatures", "model")) %>%
  filter(!(category=="function_words" & wordFeatures %in% c("socialness", "valence", "arousal", "dominance"))) %>% 
  filter(!is.na(cor))

all.res.cat.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=cor, y=jsd, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.2, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits = c(-0.16, 1), breaks = seq(0, 1, .2), expand = expansion(mult = c(.03, .08)))+
  scale_y_continuous(expand = expansion(mult = c(.05, .15)))+
  theme_bw()+
  labs(title="Nouns", x="Correlation coefficient", y="JS divergence")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

all.res.cat.long %>% subset(model=="Gpt-4o") %>% 
  ggplot(aes(x=cor, y=jsd, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.2, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits = c(-0.16, 1.05), breaks = seq(0, 1, .2), expand = expansion(mult = c(.03, .08)))+
  scale_y_continuous(expand = expansion(mult = c(.05, .15)))+
  theme_bw()+
  labs(title="Gpt-4o", x="Correlation coefficient", y="JS divergence")+
  facet_wrap(~category, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

# map_dfr(cats, function(cat_name) {
#   all.res.cat.long %>% subset(category==cat_name) %>% 
#   ggplot(aes(x=cor, y=jsd, color=wordFeatures))+
#   geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.2, linewidth=1)+ # se is 95%CI here
#   geom_point(size=2)+
#   geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
#                    box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
#                    max.overlaps = Inf, max.time = 2,
#                    segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
#   scale_x_continuous(limits = c(-0.16, 1.05), breaks = seq(0, 1, .2), expand = expansion(mult = c(.03, .08)))+
#   scale_y_continuous(expand = expansion(mult = c(.05, .15)))+
#   theme_bw()+
#   labs(title=cat_name, x="Correlation coefficient", y="JS divergence")+
#   facet_wrap(~model, strip.position="top")+
#   theme(plot.title=element_text(size=18, face="bold", color="black"),
#         axis.ticks=element_line(color="black"),
#         axis.text=element_text(size=14, color="black"),
#         axis.title=element_text(size=14, color="black"),
#         axis.text.x=element_text(size=14, color="black"),
#         strip.text=element_text(size=14),
#         legend.position="none") -> gp
#   ggsave(file=here("figures", "exp1", paste0("Scat_Cat_", cat_name, ".png")), dpi=350, width=11, height=9.5)})

map_dfr(models, function(model_name) {
  all.res.cat.long %>% subset(model==model_name) %>% 
  ggplot(aes(x=cor, y=jsd, color=wordFeatures))+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.2, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits = c(-0.16, 1.05), breaks = seq(0, 1, .2), expand = expansion(mult = c(.03, .08)))+
  scale_y_continuous(expand = expansion(mult = c(.05, .15)))+
  theme_bw()+
  labs(title=model_name, x="Correlation coefficient", y="JS divergence")+
  facet_wrap(~category, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none") -> gp
  ggsave(file=here("figures", "exp1", paste0("Scat_Cat_", model_name, ".png")), dpi=350, width=11, height=9.5)})

# Correlation coefficient
map_dfr(cats, function(cat_name){
  df <- all.res.cat.long %>% subset(category==cat_name)
  map_dfr(models, function(model_name) {
  res <- cor.test(df[df$model==model_name,]$cor, 
                  df[df$model==model_name,]$jsd, method="pearson", use="complete.obs")
  data.frame(category=cat_name, model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})})
#                category          model        cor       p.val        lwr         upr
# cor...1           other         Gpt-4o -0.1020609 0.659780952 -0.5112240  0.34481798
# cor...2           other    Gpt-4o-mini -0.4666789 0.032950728 -0.7477289 -0.04382003
# cor...3           other Llama-3.1-405B -0.3972397 0.074566625 -0.7075872  0.04157678
# cor...4           other  Llama-3.1-70B -0.4603206 0.035744408 -0.7441367 -0.03573482
# cor...5           nouns         Gpt-4o -0.4452451 0.043109748 -0.7355537 -0.01678435
# cor...6           nouns    Gpt-4o-mini -0.2274432 0.321427229 -0.6001999  0.22648033
# cor...7           nouns Llama-3.1-405B -0.3662509 0.102486254 -0.6890025  0.07772459
# cor...8           nouns  Llama-3.1-70B -0.4071896 0.066944954 -0.7134647  0.02972158
# cor...9      predicates         Gpt-4o -0.2262385 0.324066583 -0.5993867  0.22768497
# cor...10     predicates    Gpt-4o-mini -0.2119004 0.356465239 -0.5896492  0.24191551
# cor...11     predicates Llama-3.1-405B -0.3372462 0.134908079 -0.6712144  0.11053237
# cor...12     predicates  Llama-3.1-70B -0.5849936 0.005343357 -0.8116988 -0.20510206
# cor...13 function_words         Gpt-4o -0.2802953 0.231312313 -0.6430539  0.18519663
# cor...14 function_words    Gpt-4o-mini -0.3178800 0.198623006 -0.6833369  0.17495493
# cor...15 function_words Llama-3.1-405B -0.2053678 0.398977024 -0.6033012  0.27444136
# cor...16 function_words  Llama-3.1-70B -0.2826332 0.241024930 -0.6530123  0.19684615
```


```{r}
all.res.cat.long2 <- all.res.cat.long %>% pivot_longer(cols = c(cor, jsd), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = dplyr::recode(Metric, "cor" = "Correlation coefficient", "jsd" = "JS divergence"),
         category = str_replace_all(category, "_", " "),
         category = str_to_title(category),
         category = factor(category, levels = c("Nouns", "Predicates", "Function Words", "Other")))

summary.df <- all.res.cat.long2 %>% group_by(model, category, Metric) %>%
  summarise(mean = mean(Value, na.rm = TRUE), sd = sd(Value, na.rm = TRUE), se =  sd / sqrt(n()), .groups = "drop")

ggplot(all.res.cat.long2, aes(x = category, y = Value, color = wordFeatures, group = wordFeatures)) +
  geom_point(alpha = 0.6, size = 1.8, position = position_jitter(width = 0.05, height = 0)) +
  geom_line(alpha = 0.4) +
  geom_errorbar(data = summary.df,
                aes(x = category, y = mean, ymin = mean - se, ymax = mean + se),
                inherit.aes = FALSE,
                color = "black", width = 0.15, linewidth = 0.6) +
  geom_point(data = summary.df,
             aes(x = category, y = mean),
             inherit.aes = FALSE,
             color = "black", shape = 21, fill = "white", size = 2.8, stroke = 0.6) +
  facet_grid(model ~ Metric, scales = "free_y") +
  theme_bw() +
  labs(x = "", y = "Value") +
  theme(axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1),
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 13, color = "black"),
        strip.text = element_text(size = 13, color = "black", face = "bold"),
        legend.position = "none")

ggplot(all.res.cat.long2, aes(x = category, y = Value, group = wordFeatures)) +
  geom_col(data = summary.df, 
           aes(x = category, y = mean, fill = category), 
           inherit.aes = FALSE, 
           width = 0.5, alpha = 1, color = "black") +
  geom_line(aes(color = wordFeatures), alpha = 0.15, linewidth = 0.3) +
  geom_point(aes(color = wordFeatures), alpha = 0.3, size = 1.2, 
             position = position_jitter(width = 0.02, height = 0)) +
  geom_errorbar(data = summary.df, 
                aes(x = category, ymin = mean - se, ymax = mean + se),
                inherit.aes = FALSE,
                color = "black", width = 0.15, linewidth = 0.7) +
  scale_fill_manual(values = c(
    "Nouns" = "#8DD3C7",
    "Predicates" = "#FFFFB3",
    "Function Words" = "#BEBADA",
    "Other" = "#FB8072"
  )) +
  facet_grid(
    model ~ Metric,
    scales = "free_y",
    labeller = labeller(Metric = c(
      "Correlation coefficient" = "Correlation coefficient",
      "JS divergence" = "JS divergence"))) +
  scale_y_continuous(limits = c(-0.151, 1), breaks = seq(0, 1, 0.2), expand = c(0, 0.02)) +
  theme_bw() +
  labs(x = "", y = NULL) +
  theme(
    axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 13, color = "black"),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    legend.position = "none")
ggsave(file=here("figures", "exp1", "Bar_Cat.png"), dpi=350, width=6, height=8)

# Regression analysis
library(lme4)
# library(lmerTest)
library(car)
library(emmeans)

fit_cor <- lmer(cor ~ model * category + (1 | wordFeatures), data = all.res.cat.long, REML = FALSE)
# summary(fit_cor)
car::Anova(fit_cor)
emmeans(fit_cor, ~ model) %>% pairs(adjust = "holm")
emmeans(fit_cor, ~ category) %>% pairs(adjust = "holm")
emmeans(fit_cor, ~ category | model) %>% pairs(adjust = "holm")

fit_jsd <- lmer(jsd ~ model * category + (1 | wordFeatures), data = all.res.cat.long, REML = FALSE)
# summary(fit_jsd)
car::Anova(fit_jsd)
emmeans(fit_jsd, ~ model) %>% pairs(adjust = "holm")
emmeans(fit_jsd, ~ category) %>% pairs(adjust = "holm")
emmeans(fit_jsd, ~ category | model) %>% pairs(adjust = "holm")
```


# AoA (correlation)

## Overall

```{r}
aoa.human.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.human[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(Human=cor)

aoa.gpt1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.gpt1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o`=cor)

aoa.gpt2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.gpt2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Gpt-4o-mini`=cor)

aoa.meta1.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.meta1[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-405B`=cor)

aoa.meta2.res <- map_dfr(cols, function(col_name) {
  spearman_cor <- cor(d.human[["AoA"]], d.meta2[[col_name]], method="spearman", use="complete.obs")
  data.frame(wordFeatures=col_name, cor=spearman_cor)}) %>% 
  rename(`Llama-3.1-70B`=cor)

aoa.all.res <- aoa.human.res %>% left_join(aoa.gpt1.res, by="wordFeatures") %>% left_join(aoa.gpt2.res, by="wordFeatures") %>% 
  left_join(aoa.meta1.res, by="wordFeatures") %>% left_join(aoa.meta2.res, by="wordFeatures")

rm(aoa.human.res, aoa.gpt1.res, aoa.gpt2.res, aoa.meta1.res, aoa.meta2.res)
```


```{r}
aoa.all.res.long <- aoa.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

aoa.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label = number(cor, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs(cor) > 0.4), 
            size = 6, show.legend = FALSE) +
  scale_color_manual(values = c("gray20", "white")) +
  scale_fill_gradient2(limits = c(min(aoa.all.res.long$cor), max(aoa.all.res.long$cor)), breaks = seq(-1, 1, 0.2),
                       low = "#1E88E5", mid = "white", high = "#B30000", midpoint = 0,
                       name = "Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_All.png"), dpi=350, width=6, height=9)
```


```{r}
# Difference scores in correlations
aoa.all.diff <- aoa.all.res %>% mutate(across(`Gpt-4o`:`Llama-3.1-70B`, ~. -Human, .names="{col}")) %>% 
  dplyr::select(-Human)

aoa.all.diff.long <- aoa.all.diff %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         abs=abs(cor))

aoa.all.diff.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label = number(cor, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs(cor) > 0.20), 
  # geom_text(aes(label = number(abs, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs > 0.3), 
            size = 6, show.legend = FALSE) +
  scale_color_manual(values = c("gray20", "white")) +
  scale_fill_gradient2(limits = c(min(aoa.all.diff.long$cor), max(aoa.all.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits = c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low = "#1E88E5", mid = "white", high = "#B30000", midpoint = 0,
                       name = "Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_All_Diff.png"), dpi=350, width=6, height=9)
```


```{r}
## Relationships between "correlation with AoA" and "Difference in correlations between Human and Model"

all.aoa.long <- aoa.all.res.long %>% subset(model!="Human") %>% rename(corAoA=cor) %>% 
  left_join(aoa.all.diff.long, by=c("wordFeatures", "model")) %>% rename(diffCor=cor, diffCorAbs=abs) %>% 
  left_join(aoa.all.res %>% dplyr::select(wordFeatures, Human), by="wordFeatures")

all.aoa.long %>% ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
ggsave(file=here("figures", "exp1", "Scat_AoA_All.png"), dpi=350, width=11, height=9.5)   

# Correlation coefficient
map_dfr(models, function(model_name) {
  res <- cor.test(all.aoa.long[all.aoa.long$model==model_name,]$corAoA, 
                  all.aoa.long[all.aoa.long$model==model_name,]$Human, method="pearson", use="complete.obs")
  data.frame(model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})
#                  model       cor        p.val       lwr       upr
# cor...1         Gpt-4o 0.7132305 2.839279e-04 0.4067918 0.8753912
# cor...2    Gpt-4o-mini 0.6664178 9.714258e-04 0.3295320 0.8527751
# cor...3 Llama-3.1-405B 0.7314366 1.647948e-04 0.4380715 0.8839993
# cor...4  Llama-3.1-70B 0.8269629 3.793092e-06 0.6147268 0.9275313


# Regression
library(boot)
boot_fun <- function(data, indices) {
  d <- data[indices, ]
  coef(lm(Human ~ corAoA, data=d))[2]  # slope only 
  }
set.seed(123)

models
model_name <- models[1]
fit <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)
boot(all.aoa.long[all.aoa.long$model==model_name,], boot_fun, R=5000) %>% boot.ci(type="perc")

model_name <- models[2]
fit <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)
boot(all.aoa.long[all.aoa.long$model==model_name,], boot_fun, R=5000) %>% boot.ci(type="perc")

model_name <- models[3]
fit <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)
boot(all.aoa.long[all.aoa.long$model==model_name,], boot_fun, R=5000) %>% boot.ci(type="perc")

model_name <- models[4]
fit <- lm(Human ~ corAoA, data=all.aoa.long[all.aoa.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)
boot(all.aoa.long[all.aoa.long$model==model_name,], boot_fun, R=5000) %>% boot.ci(type="perc")
```


## Partial correlations

```{r}
d.freq <- read.csv("dataset/exp1/concatenated_results.csv", header=TRUE) %>% dplyr::select(seq1:AoA, frequency) %>% 
  mutate(category2=case_when(category2=="verbs" | category2=="adjectives"~"predicates", TRUE~category2)) 

d.human <- d.human %>% left_join(d.freq)
rm(d.freq)

library(ppcor)

cor(d.human$concreteness, d.human$frequency, use="complete.obs") # No multicolinearity
cols_p <- setdiff(cols, "concreteness")
# cols_p <- cols

aoa.human.partial <- map_dfr(cols_p, function(col_name) {
  tmp <- d.human[, c("AoA", col_name, "concreteness", "frequency")]
  tmp <- tmp[complete.cases(tmp), ]
  res <- pcor.test(tmp$AoA, tmp[[col_name]], tmp[, c("concreteness", "frequency")], method = "spearman")
  data.frame(wordFeatures = col_name, cor = res$estimate)}) %>% 
  rename(Human = cor)

aoa.gpt1.partial <- map_dfr(cols_p, function(col_name) {
  tmp <- data.frame(AoA = d.human$AoA, X = d.gpt1[[col_name]], concreteness = d.human$concreteness, frequency = d.human$frequency)
  tmp <- tmp[complete.cases(tmp), ]
  res <- pcor.test(tmp$AoA, tmp$X, tmp[, c("concreteness", "frequency")], method = "spearman")
  data.frame(wordFeatures = col_name, cor = res$estimate)}) %>% 
  rename(`Gpt-4o` = cor)

aoa.gpt2.partial <- map_dfr(cols_p, function(col_name) {
  tmp <- data.frame(AoA = d.human$AoA, X = d.gpt2[[col_name]], concreteness = d.human$concreteness, frequency = d.human$frequency)
  tmp <- tmp[complete.cases(tmp), ]
  res <- pcor.test(tmp$AoA, tmp$X, tmp[, c("concreteness", "frequency")], method = "spearman")
  data.frame(wordFeatures = col_name, cor = res$estimate)}) %>% 
  rename(`Gpt-4o-mini` = cor)

aoa.meta1.partial <- map_dfr(cols, function(col_name) {
  tmp <- data.frame(AoA = d.human$AoA, X = d.meta1[[col_name]], concreteness = d.human$concreteness, frequency = d.human$frequency)
  tmp <- tmp[complete.cases(tmp), ]
  res <- pcor.test(tmp$AoA, tmp$X, tmp[, c("concreteness", "frequency")], method = "spearman")
  data.frame(wordFeatures = col_name, cor = res$estimate)}) %>% 
  rename(`Llama-3.1-405B` = cor)

aoa.meta2.partial <- map_dfr(cols, function(col_name) {
  tmp <- data.frame(AoA = d.human$AoA, X = d.meta2[[col_name]], concreteness = d.human$concreteness, frequency = d.human$frequency)
  tmp <- tmp[complete.cases(tmp), ]
  res <- pcor.test(tmp$AoA, tmp$X, tmp[, c("concreteness", "frequency")], method = "spearman")
  data.frame(wordFeatures = col_name, cor = res$estimate)}) %>% 
  rename(`Llama-3.1-70B` = cor)

aoa.partial.all.res <- aoa.human.partial %>% left_join(aoa.gpt1.partial, by = "wordFeatures") %>% 
  left_join(aoa.gpt2.partial, by = "wordFeatures") %>% left_join(aoa.meta1.partial, by = "wordFeatures") %>%
  left_join(aoa.meta2.partial, by = "wordFeatures")

rm(aoa.human.partial, aoa.gpt1.partial, aoa.gpt2.partial, aoa.meta1.partial, aoa.meta2.partial)
```


```{r}
aoa.partial.all.res.long <- aoa.partial.all.res %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

aoa.partial.all.res.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label = number(cor, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs(cor) > 0.4), 
            size = 6, show.legend = FALSE) +
  scale_color_manual(values = c("gray20", "white")) +
  scale_fill_gradient2(limits = c(min(aoa.all.res.long$cor), max(aoa.all.res.long$cor)), breaks = seq(-1, 1, 0.2),
                       low = "#1E88E5", mid = "white", high = "#B30000", midpoint = 0,
                       name = "Correlation\ncoefficient\n(partial)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_Partial_All.png"), dpi=350, width=6, height=9)
```


```{r}
# Difference scores in correlations
aoa.partial.all.diff <- aoa.partial.all.res %>% mutate(across(`Gpt-4o`:`Llama-3.1-70B`, ~. -Human, .names="{col}")) %>% 
  dplyr::select(-Human)

aoa.partial.all.diff.long <- aoa.partial.all.diff %>% pivot_longer(cols=-wordFeatures, names_to="model", values_to="cor") %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         abs=abs(cor))

aoa.partial.all.diff.long %>% ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label = number(cor, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs(cor) > 0.20), 
  # geom_text(aes(label = number(abs, accuracy = 0.01, decimal.mark = ".", suffix = ""), color = abs > 0.3), 
            size = 6, show.legend = FALSE) +
  scale_color_manual(values = c("gray20", "white")) +
  scale_fill_gradient2(limits = c(min(aoa.all.diff.long$cor), max(aoa.all.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits = c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low = "#1E88E5", mid = "white", high = "#B30000", midpoint = 0,
                       name = "Difference\n(partial\ncorrelation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))
ggsave(file=here("figures", "exp1", "AoA_Partical_All_Diff.png"), dpi=350, width=6, height=9)
```


```{r}
## Relationships between "correlation with AoA" and "Difference in correlations between Human and Model"

all.aoa.partial.long <- aoa.partial.all.res.long %>% subset(model!="Human") %>% rename(corAoA=cor) %>% 
  left_join(aoa.partial.all.diff.long, by=c("wordFeatures", "model")) %>% rename(diffCor=cor, diffCorAbs=abs) %>% 
  left_join(aoa.partial.all.res %>% dplyr::select(wordFeatures, Human), by="wordFeatures")

all.aoa.partial.long %>% ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_label_repel(aes(label=wordFeatures), size=4.5, label.size=0, label.padding=unit(2.5, "pt"),
                   box.padding=unit(2, "pt"), point.padding=unit(2, "pt"), 
                   max.overlaps = Inf, max.time = 2,
                   segment.size=0.25, segment.alpha=0.8, seed=123, fill=alpha("white", 0.6))+ 
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(x="Partial correlation with AoA (Model)", y="Partial correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
ggsave(file=here("figures", "exp1", "Scat_AoA_Partial_All.png"), dpi=350, width=11, height=9.5)   

# Correlation coefficient
map_dfr(models, function(model_name) {
  res <- cor.test(all.aoa.partial.long[all.aoa.partial.long$model==model_name,]$corAoA, 
                  all.aoa.partial.long[all.aoa.partial.long$model==model_name,]$Human, method="pearson", use="complete.obs")
  data.frame(model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})
#                  model       cor        p.val       lwr       upr
# cor...1         Gpt-4o 0.7672080 7.898177e-05 0.4915821 0.9031160
# cor...2    Gpt-4o-mini 0.7438490 1.700863e-04 0.4491862 0.8925653
# cor...3 Llama-3.1-405B 0.7804068 4.922027e-05 0.5161293 0.9090058
# cor...4  Llama-3.1-70B 0.8177506 1.064736e-05 0.5880161 0.9253966

# Regression
model_name <- models[1]
fit <- lm(Human ~ corAoA, data=all.aoa.partial.long[all.aoa.partial.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)

model_name <- models[2]
fit <- lm(Human ~ corAoA, data=all.aoa.partial.long[all.aoa.partial.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)

model_name <- models[3]
fit <- lm(Human ~ corAoA, data=all.aoa.partial.long[all.aoa.partial.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)

model_name <- models[4]
fit <- lm(Human ~ corAoA, data=all.aoa.partial.long[all.aoa.partial.long$model==model_name,])
summary(fit)
car::linearHypothesis(fit, "corAoA = 1")
confint(fit)
```



## By syntactic category

```{r}
aoa.human.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.human %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(Human=cor) # errors due to impossible calculation of correlation coefficients for some features (sd=0)

aoa.gpt1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o`=cor) 

aoa.gpt2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.gpt2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Gpt-4o-mini`=cor)

aoa.meta1.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta1 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-405B`=cor)

aoa.meta2.res <- map_dfr(cats, function(cat_name) {
  df1 <- d.human %>% subset(category2==cat_name)
  df2 <- d.meta2 %>% subset(category2==cat_name)
  map_dfr(cols, function(col_name) {
  spearman_cor <- cor(df1[["AoA"]], df2[[col_name]], method="spearman", use="complete.obs")
  data.frame(category=cat_name, wordFeatures=col_name, cor=spearman_cor)})}) %>% 
  rename(`Llama-3.1-70B`=cor)

aoa.all.cat.res <- aoa.human.res %>% left_join(aoa.gpt1.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.gpt2.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.meta1.res, by=c("category", "wordFeatures")) %>% 
  left_join(aoa.meta2.res, by=c("category", "wordFeatures"))

rm(aoa.human.res, aoa.gpt1.res, aoa.gpt2.res, aoa.meta1.res, aoa.meta2.res)
```


```{r}
aoa.all.cat.res.long <- aoa.all.cat.res %>% 
  pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Human", "Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

aoa.all.cat.res.long %>% subset(category=="predicates") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.res.long$cor), max(aoa.all.cat.res.long$cor)), breaks=seq(-1,1,0.1),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  aoa.all.cat.res.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  geom_vline(xintercept=1.5, color="gray20", linewidth=1)+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.res.long$cor), max(aoa.all.cat.res.long$cor)), breaks=seq(-1,1,0.1),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Correlation\ncoefficient")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("AoA_Cat_", cat_name, ".png")), dpi=350, width=6, height=9)})
```


```{r}
# Difference scores in correlations
aoa.all.cat.diff <- aoa.all.cat.res %>% mutate(across(`Gpt-4o`:`Llama-3.1-70B`, ~. -Human, .names="{col}")) %>% 
  dplyr::select(-Human)

aoa.all.cat.diff.long <- aoa.all.cat.diff %>% pivot_longer(cols=-c(category, wordFeatures), names_to="model", values_to="cor") %>% 
  # mutate(category=factor(category, levels=c("nouns", "verbs", "adjectives", "function_words", "other"))) %>% 
  mutate(category=factor(category, levels=c("nouns", "predicates", "function_words", "other"))) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         abs=abs(cor))

aoa.all.cat.diff.long %>% subset(category=="predicates") %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  # geom_text(aes(label=number(abs, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.diff.long$cor), max(aoa.all.cat.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits=c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"))

map_dfr(cats, function(cat_name) {
  aoa.all.cat.diff.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cor))+
  # geom_tile(aes(fill=abs))+
  geom_text(aes(label=number(cor, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  # geom_text(aes(label=number(abs, accuracy=0.01, decimal.mark=".", suffix="")), size=6, color="gray30")+
  scale_fill_gradient2(limits=c(min(aoa.all.cat.diff.long$cor), max(aoa.all.cat.diff.long$cor)), breaks=seq(-3,3,0.1),
  # scale_fill_gradient2(limits=c(min(aoa.all.diff.long$abs), max(aoa.all.diff.long$abs)), breaks=seq(-1,1,0.05),
                       low="#00A0E9", mid="white", high="#E60012", midpoint=0, 
                       name="Difference\n(correlation)")+
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  facet_wrap(~category, strip.position="bottom")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black")) -> gp
  ggsave(file=here("figures", "exp1", paste0("AoA_Cat_Diff_", cat_name, ".png")), dpi=350, width=6, height=9)})
```


```{r}
## Relationships between "correlation with AoA" and "Difference in correlations between Human and Model"

all.aoa.cat.long <- aoa.all.cat.res.long %>% subset(model!="Human") %>% rename(corAoA=cor) %>% 
  left_join(aoa.all.cat.diff.long, by=c("category", "wordFeatures", "model")) %>% rename(diffCor=cor, diffCorAbs=abs) %>% 
  left_join(aoa.all.cat.res %>% dplyr::select(category, wordFeatures, Human), by=c("category", "wordFeatures"))

all.aoa.cat.long %>% subset(category=="nouns") %>% 
  ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(title="Nouns", x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

map_dfr(cats, function(cat_name) {
  all.aoa.cat.long %>% subset(category==cat_name) %>% 
  ggplot(aes(x=corAoA, y=Human, color=wordFeatures))+
  geom_abline(slope=1, intercept=0, color="black", linetype="dashed", size=0.8)+
  geom_smooth(color="gray60", method="lm", se=TRUE, alpha=0.3, linewidth=1)+ # se is 95%CI here
  geom_point(size=2)+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  scale_x_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  scale_y_continuous(limits=c(-0.6,0.6), breaks=seq(-1,1,0.2))+
  theme_bw()+
  labs(title=cat_name, x="Correlation with AoA (Model)", y="Correlation with AoA (Human)")+
  facet_wrap(~model, strip.position="top")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")
  ggsave(file=here("figures", "exp1", paste0("Scat_AoA_Cat_", cat_name, ".png")), dpi=350, width=11, height=9.5)})

# Correlation coefficient
map_dfr(cats, function(cat_name){
  df <- all.aoa.cat.long %>% subset(category==cat_name)
  map_dfr(models, function(model_name) {
    res <- cor.test(df[df$model==model_name,]$corAoA, 
                    df[df$model==model_name,]$Human, method="pearson", use="complete.obs")
    data.frame(category=cat_name, model=model_name, cor=res$estimate, p.val=res$p.value, lwr=res$conf.int[1], upr=res$conf.int[2])})})
#                category          model         cor        p.val         lwr       upr
# cor...1           other         Gpt-4o 0.786278904 2.378628e-05  0.53679411 0.9093190
# cor...2           other    Gpt-4o-mini 0.604008118 3.735852e-03  0.23310051 0.8214961
# cor...3           other Llama-3.1-405B 0.799343586 1.380247e-05  0.56136472 0.9152193
# cor...4           other  Llama-3.1-70B 0.718636392 2.426169e-04  0.41600516 0.8779579
# cor...5           nouns         Gpt-4o 0.660693240 1.113045e-03  0.32038399 0.8499606
# cor...6           nouns    Gpt-4o-mini 0.614619297 3.029881e-03  0.24899721 0.8269083
# cor...7           nouns Llama-3.1-405B 0.487309967 2.504885e-02  0.07044167 0.7592722
# cor...8           nouns  Llama-3.1-70B 0.683900484 6.295373e-04  0.35786709 0.8613038
# cor...9      predicates         Gpt-4o 0.695015143 4.705765e-04  0.37619888 0.8666746
# cor...10     predicates    Gpt-4o-mini 0.753171369 8.116992e-05  0.47636812 0.8941421
# cor...11     predicates Llama-3.1-405B 0.641691271 1.715049e-03  0.29046652 0.8405403
# cor...12     predicates  Llama-3.1-70B 0.749616350 9.157450e-05  0.47003152 0.8924929
# cor...13 function_words         Gpt-4o 0.007035311 9.771960e-01 -0.44860750 0.4597754
# cor...14 function_words    Gpt-4o-mini 0.202914138 4.193551e-01 -0.29157849 0.6118234
# cor...15 function_words Llama-3.1-405B 0.088441158 7.271109e-01 -0.39472742 0.5332913
# cor...16 function_words  Llama-3.1-70B 0.095281520 7.068563e-01 -0.38888788 0.5382098
```



# Clustering

## tSNE (Overall)

```{r}
d.all.int <- cor.all.res.long %>% left_join(jsd.all.res.long, by=c("wordFeatures", "model")) %>% 
  left_join(all.aoa.long, by=c("wordFeatures", "model")) %>%
  # left_join(all.aoa.partial.long, by=c("wordFeatures", "model")) %>%
  mutate(corAoAR = corAoA / Human) %>% 
  dplyr::select(wordFeatures, model, cor, jsd, corAoA, diffCorAbs) # alternatives: diffCor, diffCorAbs, corAoAR

# tSNE
set.seed(123)
# ?Rtsne
tsne.out <- Rtsne(d.all.int %>% dplyr::select(-wordFeatures, -model), pca=TRUE, perplexity=6, theta=0)

d.all.int.res <- d.all.int %>% cbind(as.data.frame(tsne.out$Y)) %>% 
  mutate(wordFeatures=factor(wordFeatures, 
                             levels=c("concreteness", "imageability", "adultBOI", "childBOI", "iconicity", "socialness", "babiness",
                                      "valence", "arousal", "dominance",
                                      "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual", "Foot_leg", "Hand_arm", 
                                      "Head", "Mouth", "Torso")),
         model=factor(model, levels=c("Gpt-4o", "Gpt-4o-mini", 
                                      "Llama-3.1-405B", "Llama-3.1-70B")))

d.all.int.res %>% ggplot(aes(x=V1, y=V2, color=wordFeatures))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  geom_point(size=4)+
  geom_hline(yintercept=20, linetype="dashed")+
  geom_hline(yintercept=0, linetype="dashed")+
  geom_hline(yintercept=-25, linetype="dashed")+
  facet_grid(~model)+
  scale_y_continuous(breaks=seq(-50,50,10))+
  scale_x_continuous(breaks=seq(-50,50,10))+
  theme_bw()+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none")

# K-means
k.max <- 10
?kmeans
wss <- sapply(1:k.max, function(k){kmeans(tsne.out$Y, k, iter.max=15, nstart=50)$tot.withinss})
plot(1:k.max, wss, type="b", pch=19, frame=FALSE, xlab="#clusters K", ylab="Total within-clusters sum of squares")

km <- kmeans(tsne.out$Y, 4, iter.max=15, nstart=50)

d.all.int.res <- d.all.int.res %>% cbind(data.frame(cluster=as.factor(km$cluster))) %>% 
  mutate(cluster=case_match(cluster, "1"~"C", "2"~"D", "3"~"A", "4"~"B"))
  # mutate(cluster=case_match(cluster, "1"~"D", "2"~"C", "3"~"A"))

d.all.int.res %>% ggplot(aes(x=V1, y=V2, color=cluster, shape=cluster))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf, show.legend=FALSE)+
  geom_point(size=4, alpha=0.6)+
  # geom_hline(yintercept=20, linetype="dashed")+
  # geom_hline(yintercept=0, linetype="dashed")+
  # geom_hline(yintercept=-25, linetype="dashed")+
  # scale_color_manual(values = c("A" = "#81C784", "B" = "#C8E6C9",  "C" = "#FFF59D", "D" = "#EF9A9A")) +
  facet_wrap(~model)+
  scale_y_continuous(breaks=seq(-50,50,10))+
  scale_x_continuous(breaks=seq(-50,50,10))+
  theme_bw()+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14))
ggsave(file=here("figures", "exp1", "tsne_All.png"), dpi=350, width=8, height=10)

d.all.int.res.long <- d.all.int.res %>% 
  # mutate(group=case_when(V2< -25~"A", V2<0~"B", V2<20~"C", TRUE~"D")) %>% 
  pivot_longer(c(cor:diffCorAbs), names_to="index", values_to="value") %>% 
  mutate(index=case_match(index, "cor"~"Correlation\n(human data)", "jsd"~"JS divergence\n(human data)",
                          "corAoA"~"Correlation\n(AoA)", "diffCorAbs"~"Difference in\ncorrelation (AoA)"), 
         index=factor(index, levels=c("Correlation\n(human data)", "JS divergence\n(human data)", 
                                      "Correlation\n(AoA)", "Difference in\ncorrelation (AoA)")))

d.all.int.res.long %>% 
  mutate(model=factor(model, levels=c("Gpt-4o",  "Gpt-4o-mini",
                     "Llama-3.1-405B", "Llama-3.1-70B"))) %>% 
  ggplot(aes(x=cluster, y=value, color=wordFeatures))+
  geom_jitter(size=2, height=0, width=0.05, alpha=0.5)+
  geom_boxplot(fill="white", color="gray20", alpha=0.5)+
  # geom_hline(yintercept=0, linetype="dashed")+
  facet_grid(index~model, scales="free")+
  theme_bw()+
  labs(y="Value", x="Cluster")+
  theme(plot.title=element_text(size=18, face="bold", color="black"),
     axis.ticks=element_line(color="black"),
     axis.text=element_text(size=14, color="black"),
     axis.title=element_text(size=14, color="black"),
     axis.text.x=element_text(size=14, color="black"),
     strip.text=element_text(size=14),
     legend.position="none") 
# ggsave(file=here("figures", "exp1", "tsne_box_All.png"), dpi=350, width=8, height=7.5)


summary.df <- d.all.int.res.long %>% group_by(model, index, cluster) %>%
  summarise(mean = mean(value, na.rm = TRUE), se = sd(value, na.rm = TRUE) / sqrt(n()), .groups = "drop")

d.all.int.res.long %>% mutate(model = factor(model, levels = c("Gpt-4o", "Gpt-4o-mini", "Llama-3.1-405B", "Llama-3.1-70B"))) %>%
  ggplot(aes(x = cluster, y = value, group = wordFeatures)) +
  geom_col(data = summary.df, aes(x = cluster, y = mean, fill = cluster),
           inherit.aes = FALSE, width = 0.55, color = "black", alpha = 0.9) +
  geom_errorbar(data = summary.df,
                aes(x = cluster, ymin = mean - se, ymax = mean + se),
                inherit.aes = FALSE,
                width = 0.15, linewidth = 0.7, color = "black") +
  geom_point(aes(color = wordFeatures), alpha = 0.3, size = 1.2,
             position = position_jitter(width = 0.05, height = 0)) +
  scale_fill_manual(values = c("A" = "#81C784", "B" = "#C8E6C9",  "C" = "#FFF59D", "D" = "#EF9A9A"), name = "Cluster") +
  facet_grid(index ~ model, scales = "free_y") +
  theme_bw() +
  labs(x = "Cluster", y = "Value") +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "black"),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 13, color = "black"),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    legend.position = "none")
ggsave(file=here("figures", "exp1", "tsne_box_All.png"), dpi=350, width=8, height=7.5)


d.all.int.res.long %>% dplyr::select(wordFeatures, model, cluster) %>% 
  unique() %>% 
  ggplot(aes(x=model, y=fct_rev(wordFeatures)))+
  geom_tile(aes(fill=cluster))+
  geom_text(aes(label=cluster), size=6, color="gray30")+
  # scale_fill_brewer(palette="BuGn", name="Cluster")+
  scale_fill_manual(values = c("A" = "#81C784", "B" = "#C8E6C9",  "C" = "#FFF59D", "D" = "#EF9A9A"), name = "Cluster") +
  scale_x_discrete(position="top")+
  theme_bw()+
  labs(x="", y="Word features")+
  theme(axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black", angle=45, hjust=0),
        strip.text=element_text(size=14),
        legend.title=element_text(size=14, color="black", hjust=0.2),
        legend.text=element_text(size=12, color="black"),
        plot.margin=margin(0, 40, 0, 0)) # (Top, Right, Bottom, Left)
ggsave(file=here("figures", "exp1", "tsne_group_All.png"), dpi=350, width=6, height=9)
```


# PCA + GMM

```{r}
library(mclust)
library(cluster)
library(flexclust)

# ---- PCA -> GMM ----
set.seed(123)
pca.out <- prcomp(d.all.int %>% dplyr::select(-wordFeatures, -model), scale. = TRUE)

expl.var <- cumsum(pca.out$sdev^2 / sum(pca.out$sdev^2))
n.pc <- which(expl.var >= 0.95)[1]
pca.data <- as.data.frame(pca.out$x[, 1:n.pc])

# Gaussian Mixture Model (GMM)
gmm.out <- Mclust(pca.data, G = 4)  # Num of group = 4

d.all.int.gmm <- d.all.int %>% mutate(cluster_gmm = as.factor(gmm.out$classification))

# ---- Agreement: tSNE+K-means vs. PCA+GMM  ----
d.compare <- d.all.int.res %>% mutate(cluster_kmeans = as.factor(km$cluster)) %>%
  left_join(d.all.int.gmm %>% dplyr::select(model, wordFeatures, cluster_gmm), by = c("model", "wordFeatures"))

# ARI (Adjusted Rand Index)
ari_value <- randIndex(d.compare$cluster_kmeans, d.compare$cluster_gmm, correct = TRUE)
print(paste("Adjusted Rand Index (t-SNE+k-means vs PCA+GMM):", round(ari_value, 3)))

# ---- Silhouette (optional) ----
sil <- silhouette(as.numeric(d.all.int.gmm$cluster_gmm), dist(pca.data))
mean_sil <- mean(sil[, 3])
print(paste("Mean silhouette width (PCA+GMM):", round(mean_sil, 3)))
```


# UMAP + GMM

```{r}
# ---- UMAP -> GMM ----
library(umap)
set.seed(123)
umap.out <- umap(d.all.int %>% dplyr::select(-wordFeatures, -model))
gmm.umap <- Mclust(umap.out$layout, G = 4)
d.all.int.umap <- d.all.int %>% mutate(cluster_gmm = as.factor(gmm.umap$classification))

# ---- Agreement: tSNE+K-means vs. UMAP+GMM  ----
d.compare <- d.all.int.res %>% mutate(cluster_kmeans = as.factor(km$cluster)) %>%
  left_join(d.all.int.umap %>% dplyr::select(model, wordFeatures, cluster_gmm), by = c("model", "wordFeatures"))

# Adjusted Rand Index
ari_value <- randIndex(d.compare$cluster_kmeans, d.compare$cluster_gmm, correct = TRUE)
print(paste("Adjusted Rand Index (t-SNE+k-means vs UMAP+GMM):", round(ari_value, 2)))

# ---- Silhouette (optional) ----
sil <- silhouette(as.numeric(d.all.int.umap$cluster_gmm),
                  dist(umap.out$layout))
mean_sil <- mean(sil[, 3])
print(paste("Mean silhouette width (UMAP+GMM):", round(mean_sil, 3)))


# ---- Visualization ----

d.all.int.umap <- d.all.int %>% cbind(as.data.frame(umap.out$layout)) %>%
  mutate(cluster = as.factor(gmm.umap$classification),
         wordFeatures = factor(wordFeatures, levels = c("concreteness", "imageability", "adultBOI", "childBOI",
          "iconicity", "socialness", "babiness", "valence", "arousal", "dominance",
          "Auditory", "Gustatory", "Haptic", "Interoceptive", "Olfactory", "Visual",
          "Foot_leg", "Hand_arm", "Head", "Mouth", "Torso")),
    model = factor(model, levels = c("Gpt-4o", "Gpt-4o-mini", "Llama-3.1-405B", "Llama-3.1-70B")),
    cluster = case_match(cluster,"1" ~ "A", "2" ~ "D", "3" ~ "C", "4" ~ "B"))

d.all.int.umap %>% ggplot(aes(x = V1, y = V2, color = cluster, shape = cluster)) +
  geom_text_repel(aes(label = wordFeatures), size = 5, max.overlaps = Inf, show.legend = FALSE) +
  geom_point(size = 4, alpha = 0.6) +
  facet_wrap(~model) +
  scale_y_continuous(breaks = seq(-8, 8, 2)) +
  scale_x_continuous(breaks = seq(-8, 8, 2)) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14))
# ggsave(file = here("figures", "exp1", "umap_All.png"), dpi = 350, width = 8, height = 10)

d.all.int.umap.long <- d.all.int.umap %>% pivot_longer(c(cor:diffCorAbs), names_to = "index", values_to = "value") %>%
  mutate(index = case_match(index,
      "cor" ~ "Correlation\n(human data)",
      "jsd" ~ "JS divergence\n(human data)",
      "corAoA" ~ "Correlation\n(AoA)",
      "diffCorAbs" ~ "Difference in\ncorrelation (AoA)"),
    index = factor(index,
      levels = c(
        "Correlation\n(human data)",
        "JS divergence\n(human data)",
        "Correlation\n(AoA)",
        "Difference in\ncorrelation (AoA)")))

summary.df.umap <- d.all.int.umap.long %>% group_by(model, index, cluster) %>%
  summarise(mean = mean(value, na.rm = TRUE), se = sd(value, na.rm = TRUE) / sqrt(n()), .groups = "drop")

d.all.int.umap.long %>% mutate(model = factor(model, levels = c("Gpt-4o", "Gpt-4o-mini", "Llama-3.1-405B", "Llama-3.1-70B"))) %>%
  ggplot(aes(x = cluster, y = value, group = wordFeatures)) +
  geom_col(data = summary.df.umap, aes(x = cluster, y = mean, fill = cluster),
    inherit.aes = FALSE, width = 0.55, color = "black", alpha = 0.9) +
  geom_errorbar(data = summary.df.umap, aes(x = cluster, ymin = mean - se, ymax = mean + se),
    inherit.aes = FALSE,
    width = 0.15, linewidth = 0.7, color = "black"
  ) +
  geom_point(aes(color = wordFeatures), alpha = 0.3, size = 1.2, position = position_jitter(width = 0.05, height = 0)) +
  scale_fill_manual(values = c("A" = "#81C784", "B" = "#C8E6C9",  "C" = "#FFF59D", "D" = "#EF9A9A"), name = "Cluster") +
  facet_grid(index ~ model, scales = "free_y") +
  theme_bw() +
  labs(x = "Cluster", y = "Value") +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "black"),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 13, color = "black"),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    legend.position = "none")
ggsave(file = here("figures", "exp1", "umap_box_All.png"), dpi = 350, width = 8, height = 7.5)

d.all.int.umap.long %>% dplyr::select(wordFeatures, model, cluster) %>%
  unique() %>%
  ggplot(aes(x = model, y = fct_rev(wordFeatures))) +
  geom_tile(aes(fill = cluster)) +
  geom_text(aes(label = cluster), size = 6, color = "gray30") +
  scale_fill_manual(values = c("A" = "#81C784", "B" = "#C8E6C9",  "C" = "#FFF59D", "D" = "#EF9A9A"), name = "Cluster") +
  scale_x_discrete(position = "top") +
  theme_bw() +
  labs(x = "", y = "Word features") +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black", angle = 45, hjust = 0),
    strip.text = element_text(size = 14),
    legend.title = element_text(size = 14, color = "black", hjust = 0.2),
    legend.text = element_text(size = 12, color = "black"),
    plot.margin = margin(0, 40, 0, 0))
ggsave(file = here("figures", "exp1", "umap_group_All.png"), dpi = 350, width = 6, height = 9)
```


## PCA (Overall)

```{r}
pca <- prcomp(d.all.int %>% dplyr::select(-wordFeatures, -model), scale=TRUE)

pca.cont <- data.frame(t(summary(pca)$importance)) %>% 
  tibble::rownames_to_column(var="PCs") %>% 
  tibble::rownames_to_column(var="PCs2") %>% 
  mutate(PCs2 = as.numeric(PCs2))

pca.cont %>% ggplot(aes(x=PCs2, y=Proportion.of.Variance))+
  geom_bar(stat="identity")+
  geom_line(aes(y=Cumulative.Proportion), color="blue", linewidth=3)+
  geom_abline(slope=0, intercept=0.1, color="red")+
  theme_bw()+
  scale_x_continuous(breaks=1:20)+
  labs(x="Principal components", y="Proportion of variance")+
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16))

pca.axis <- data.frame(pca$rotation) %>% tibble::rownames_to_column(var="Column")

library(ggsci)
pca.axis %>% ggplot(aes(x=PC1, y=PC2))+
  #geom_point()+
  geom_text(aes(label=Column), hjust=0, vjust=0)+
  geom_segment(aes(x=0, y=0, xend=PC1, yend=PC2),linewidth=.4, arrow=arrow(length=unit(.1, 'cm')))+
  theme_bw()+
  scale_colour_d3()+ 
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),  
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16))

d.all.int %>% cbind(data.frame(pca$x)) %>% 
  ggplot(aes(x=PC1, y=PC2, color=wordFeatures))+
  geom_text_repel(aes(label=wordFeatures), size=5, max.overlaps=Inf)+
  geom_point(size=4)+
  facet_grid(~model)+
  theme_bw()+
   theme(plot.title=element_text(size=18, face="bold", color="black"),
        axis.ticks=element_line(color="black"),
        axis.text=element_text(size=14, color="black"),
        axis.title=element_text(size=14, color="black"),
        axis.text.x=element_text(size=14, color="black"),
        strip.text=element_text(size=14),
        legend.position="none") 
```



# Other exploratory analyses

## Prediction of AoA (Overall)

```{r}
# Human
d.human.aoa <- d.human %>% dplyr::select(-socialness, -imageability) %>%  # socialness and imageability had >200 missing items
  filter(complete.cases(dplyr::select(., AoA:babiness))) %>% 
  mutate(across(concreteness:babiness, ~ scale(.)[,1], .names="{.col}")) %>% 
  mutate(Human="yes")

m.human.null <- lm(AoA ~ 1, data=d.human.aoa %>% dplyr::select(AoA, concreteness:babiness))
m.human.full <- lm(AoA ~ ., data=d.human.aoa %>% dplyr::select(AoA, concreteness:babiness))
anova(m.human.null, m.human.full)

summary(m.human.full)
vif(m.human.full)
```

```{r}
# gpt1
d.gpt1.aoa <- d.gpt1 %>% dplyr::select(-socialness, -imageability) %>%
  mutate(across(concreteness:babiness, ~ scale(.)[,1], .names="{.col}")) %>% 
  left_join(d.human %>% dplyr::select(seq1, AoA), by="seq1") %>% 
  left_join(d.human.aoa %>% dplyr::select(seq1, Human), by="seq1") %>% 
  mutate(Human=case_when(Human=="yes"~"yes", TRUE~"no")) %>% 
  filter(complete.cases(dplyr::select(., concreteness:AoA))) %>% 
  dplyr::select(-adultBOI) %>% 
  subset(Human=="yes")

m.gpt1.null <- lm(AoA ~ 1, data=d.gpt1.aoa %>% dplyr::select(AoA, concreteness:babiness))
m.gpt1.full <- lm(AoA ~ ., data=d.gpt1.aoa %>% dplyr::select(AoA, concreteness:babiness))
anova(m.gpt1.null, m.gpt1.full)

summary(m.gpt1.full)
vif(m.gpt1.full)

m.gpt1.sel <- step(m.gpt1.null, scope=list(lower=m.gpt1.null, upper=m.gpt1.full), direction="both", k=log(nrow(d.gpt1.aoa)))
summary(m.gpt1.sel)
vif(m.gpt1.sel)
```

